---
title: "Additional analysis"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
    code_folding: hide
  word_document:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library, echo = FALSE}
library(tidyverse)
library(gtsummary)
library(survminer)
library(survival)
theme_gtsummary_compact()
theme_set(theme_classic())
```

```{r load}
data <- 
  readxl::read_xlsx(paste0(here::here(), 
                           "/sample_status_20250613.xlsx"))
sequenced_patient_data <- read_rds(paste0(
  here::here(),
  "/processed data",
  "/Identified breast data_2025-08-15.rds"))

drug_tertile <- read_rds(paste0(
  here::here(),
  "/processed data",
  "/Cumulative dose score_2025-05-06.rds"))
```

```{r merge}
data_merge <- data %>% 
  mutate(deidentified_patient_id = paste0("breast_study_", sampleID)) %>% 
  left_join(., sequenced_patient_data %>% 
              distinct(deidentified_patient_id, .keep_all = TRUE),
            by = join_by(deidentified_patient_id)) %>% 
  left_join(., drug_tertile %>% 
              select(deidentified_patient_id, tertile_patient_pre_firstseqsample) %>%
              filter(!is.na(tertile_patient_pre_firstseqsample)) %>% 
              distinct(deidentified_patient_id, .keep_all = TRUE),
            by = join_by(deidentified_patient_id))
```

```{r cleaning}
data_merge <- data_merge %>% 
  mutate(tnm_stage = case_when(
    str_detect(tnm_stage, "0")       ~ "0",
    str_detect(tnm_stage, "1")       ~ "1",
    str_detect(tnm_stage, "2")       ~ "2",
    str_detect(tnm_stage, "3")       ~ "3",
    str_detect(tnm_stage, "4")       ~ "4",
    str_detect(tnm_stage, "99")      ~ NA_character_,
    TRUE                             ~ tnm_stage
  )) %>% 
  mutate(tnm_stage_cat = case_when(
    tnm_stage == "0" |
    tnm_stage == "1" |
    tnm_stage == "2"                 ~ "Early stage",
    tnm_stage == "3" |
    tnm_stage == "4"                 ~ "Late stage",
    TRUE                             ~ tnm_stage
  ), tnm_stage_cat = factor(tnm_stage_cat, levels = c("Early stage", "Late stage"))) %>% 
  mutate_at(c("er", "pr", "her2"), ~ factor(., levels = c("neg", "pos"))) %>% 
  mutate(status_CH = factor(status_CH, levels = c("noCH", "CH"))) %>% 
  mutate(cond = case_when(
    status_selection == "NegativeSelection"      ~ "Neg/Not impacted",
    status_selection == "NotChanged"             ~ "Neg/Not impacted",
    status_selection == "PositiveSelection"      ~ "Pos"
  ), cond = factor(cond, levels = c("Neg/Not impacted", "Pos"))) %>% 
  
  mutate(her2 = case_when(
    str_detect(deidentified_patient_id, "007422")      ~ "neg",
    TRUE                                               ~ her2
  ))
  
# write_rds(data_merge %>% 
#             select(deidentified_patient_id, 
#                    age_at_sample, tnm_stage, tnm_stage_cat,
#                    er, pr, her2, status_CH, status_selection, cond, 
#                    os_event, os_time_from_tx_start_months, pfs_event, pfs_time_months), 
#           "data_meerge_updated.rds")
```

# I. OS Overall population
## 1. Summary 
```{r Clinical table}
data_merge %>% 
  select(age_at_sample, tnm_stage, tnm_stage_cat,
         er, pr, her2, status_CH, status_selection, cond,
         tertile_patient_pre_firstseqsample) %>% 
  mutate(age_cat = case_when(
    age_at_sample < 40         ~ "20-39",
    age_at_sample < 50          ~ "40-49",
    age_at_sample < 60          ~ "50-59",
    age_at_sample >= 60          ~ "60-80",
  )) %>% 
  tbl_summary() %>% 
  bold_labels()
```

## 2. Hazard ratios and 95% confidence intervals
```{r}
tbl1 <- tbl_uvregression(
  data_merge,
  method = coxph,
  y = Surv(time = os_time_from_tx_start_months, event = os_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", #"status_CH",
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = data_merge$os_time_from_tx_start_months,
                   event = data_merge$os_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + #status_CH +
                tertile_patient_pre_firstseqsample,
              data = data_merge)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>% 
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

```{r}
tbl1 <- tbl_uvregression(
  data_merge,
  method = coxph,
  y = Surv(time = os_time_from_tx_start_months, event = os_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", "cond",
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = data_merge$os_time_from_tx_start_months,
                   event = data_merge$os_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + cond +
                tertile_patient_pre_firstseqsample,
              data = data_merge)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>% 
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

## 3. KM
```{r Survival, fig.height = 7}
ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ tertile_patient_pre_firstseqsample,
                   data=data_merge),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

# II. OS Non-CH subset
```{r}
non_ch_subset <- data_merge %>% 
  mutate(status_selection = factor(status_selection)) %>% 
  filter(status_CH == "noCH")
```

## 1. Summary 
```{r Clinical table non_ch_subset}
non_ch_subset %>% 
  select(age_at_sample, tnm_stage, tnm_stage_cat, 
         er, pr, her2, status_CH, status_selection, cond,
         tertile_patient_pre_firstseqsample) %>% 
  tbl_summary() %>% 
  bold_labels()
```

## 2. Hazard ratios and 95% confidence intervals
```{r HR non_ch_subset}
tbl1 <- tbl_uvregression(
  non_ch_subset,
  method = coxph,
  y = Surv(time = os_time_from_tx_start_months, event = os_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", #"cond",
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = non_ch_subset$os_time_from_tx_start_months,
                   event = non_ch_subset$os_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + #cond +
                tertile_patient_pre_firstseqsample,
              data = non_ch_subset)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

## 3. KM
```{r Survival non_ch_subset, fig.height = 7}
ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ tertile_patient_pre_firstseqsample,
                   data=non_ch_subset),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

# III. PFS Overall population

## 1. Hazard ratios and 95% confidence intervals
```{r PFS}
tbl1 <- tbl_uvregression(
  data_merge,
  method = coxph,
  y = Surv(time = pfs_time_months, event = pfs_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", #"status_CH", 
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = data_merge$pfs_time_months,
                   event = data_merge$pfs_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + #status_CH +
                tertile_patient_pre_firstseqsample,
              data = data_merge)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

```{r}
tbl1 <- tbl_uvregression(
  data_merge,
  method = coxph,
  y = Surv(time = pfs_time_months, event = pfs_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", "cond", 
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = data_merge$pfs_time_months,
                   event = data_merge$pfs_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + cond +
                tertile_patient_pre_firstseqsample,
              data = data_merge)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

## 2. KM
The KM are not ajusted
```{r Survival PFS, fig.height = 7}
ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ tertile_patient_pre_firstseqsample,
                   data=data_merge),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

# IV. PFS Non-CH subset

## 1. Hazard ratios and 95% confidence intervals
```{r HR PFS non_ch_subset}
tbl1 <- tbl_uvregression(
  non_ch_subset,
  method = coxph,
  y = Surv(time = pfs_time_months, event = pfs_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", #"cond",
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = non_ch_subset$pfs_time_months,
                   event = non_ch_subset$pfs_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + #cond +
              tertile_patient_pre_firstseqsample,
              data = non_ch_subset)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

## 2. KM
```{r PFS non_ch_subset, fig.height = 7}
ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ tertile_patient_pre_firstseqsample,
                   data=non_ch_subset),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

# V. Selection stratification - KM OS
## 1. Neg/Not impacted selection
```{r}
neg_selection <- data_merge %>% 
  filter(cond == "Neg/Not impacted")

ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ tertile_patient_pre_firstseqsample,
                   data=neg_selection),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```

## 2. Positive selection
```{r}
pos_selection <- data_merge %>% 
  filter(cond == "Pos")

ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ tertile_patient_pre_firstseqsample,
                   data=pos_selection),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

# VI. Selection stratification - KM PFS
## 1. Neg/Not impacted selection
```{r}
neg_selection <- data_merge %>% 
  filter(cond == "Neg/Not impacted")

ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ tertile_patient_pre_firstseqsample,
                   data=neg_selection),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```

## 2. Positive selection
```{r}
pos_selection <- data_merge %>% 
  filter(cond == "Pos")

ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ tertile_patient_pre_firstseqsample,
                   data=pos_selection),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

# VII. Investigate homogenous patient group
```{r redo drug class but keeping drug name in it}
path_raw <- fs::path("", "Volumes", "Lab_Gillis", "Data", "Breast",
                     "Breast_R01", "sequential_samples_hossein")
path_raw <- fs::path("", "Volumes", "Gillis_Research", "Lab_Data", "BreastCHEvolution",
                     "ProcessedData")
drug_dose <- 
  readxl::read_xlsx(paste0(
    path_raw, 
    "/BreastCH_DosingChemo_20250402.xlsx"),
    sheet = "Normalized") %>% 
  janitor::clean_names()

sample_dates <- 
  readxl::read_xlsx(paste0(
    path_raw, 
    "/BreastCH_DosingChemo_20250402.xlsx"),
    sheet = "Chemo_patient_MRN_chemotherapy_") %>% 
  janitor::clean_names()

drug_class <- 
  readxl::read_xlsx(paste0(
    path_raw, 
    "/BreastCH_BoltonChemoDosingNatGenet2020.xlsx"),
    sheet = "Sup. Table 1", skip = 1) %>% 
  janitor::clean_names()

path_save <- fs::path("", "Volumes", "Gillis_Research","Christelle Colin-Leitzinger", 
                      "Breast CH", "sequential_samples_hossein")

sequenced_patient_data <- 
  read_csv(paste0(# path_save, 
    here::here(), 
    "/processed data",
    "/Identified breast data with sequenced sequential sample and clinical-CBC-Cytopenia-NADIR_2025-03-31.csv"))


############################################################ II ### Cumulative doses----
sample_dates <- sample_dates %>% 
  select(mrn, pre_sample_collection_date, last_post_sample_collection_date) %>% 
  distinct(mrn, .keep_all = TRUE) %>% 
  mutate_at(c("pre_sample_collection_date", "last_post_sample_collection_date"), ~ as.Date(., origin = "1899-12-30")) %>% 
  left_join(., sequenced_patient_data %>% 
              # select(mrn, treatment_type, seqsample_date_chemo, seqsample_date_rad, 
              #        seqsample_date_Achemo, seqsample_date_hormone) %>% 
              group_by(mrn) %>% 
              fill(seqsample_date_chemo, seqsample_date_rad, 
                   seqsample_date_Achemo, seqsample_date_hormone, .direction = "updown") %>% 
              distinct(mrn, .keep_all = TRUE) %>% 
              mutate(date_of_first_seq_sample = case_when(
                treatment_type == "chemo"              ~ seqsample_date_chemo,
                treatment_type == "chemorad"           ~ seqsample_date_Achemo,
                treatment_type == "hormone"            ~ seqsample_date_hormone,
                treatment_type == "radiation"          ~ seqsample_date_rad
              )) %>% 
              select(mrn, date_of_first_seq_sample),
            by = "mrn") %>% 
  mutate(mrn = as.character(mrn))
  
drug_dose1 <- drug_dose %>% 
  mutate(mrn = as.character(mrn)) %>% 
  left_join(., sample_dates, by = "mrn") %>% 
  mutate(pre_to_first_sample = case_when(
    date >= pre_sample_collection_date &
      date <= date_of_first_seq_sample                  ~ "Yes"
  )) %>% 
  mutate(pre_to_last_sample = case_when(
    date >= pre_sample_collection_date &
      date <= last_post_sample_collection_date          ~ "Yes"
  )) %>% 
  # fix drug name and add class
  mutate(chemotherapy_drug = case_when(
    chemotherapy_drug == "5 fu"                         ~ "fluorouracil",
    chemotherapy_drug == "adriamycin"                   ~ "doxorubicin",
    chemotherapy_drug == "abraxane (form of taxol)"     ~ "abraxane",
    TRUE                                                ~ chemotherapy_drug
  )) %>% 
  # add 
  left_join(., drug_class, 
            by = c("chemotherapy_drug" = "drug_name")) %>% 
  # calculate cumulative dose per patient per drug
  group_by(mrn, chemotherapy_drug, pre_to_first_sample) %>% 
  mutate(cumulative_dose_pre_firstseqsample = case_when(
    pre_to_first_sample == "Yes"                        ~ sum(dose_mg_per_m2)
  )) %>% 
  group_by(mrn, chemotherapy_drug, pre_to_last_sample) %>% 
  mutate(cumulative_dose_pre_lastseqsample = case_when(
    pre_to_last_sample == "Yes"                         ~ sum(dose_mg_per_m2)
  )) %>% 
  ungroup() %>% 
  select(mrn, chemotherapy_drug, narrow_drug_class_cytotoxic_only, 
         cumulative_dose_pre_firstseqsample, cumulative_dose_pre_lastseqsample) %>% 
  arrange(mrn, chemotherapy_drug, 
          cumulative_dose_pre_firstseqsample, cumulative_dose_pre_lastseqsample) %>% 
  distinct(mrn, chemotherapy_drug, .keep_all = TRUE) %>% 
  # create tertile categories
  # group_by(chemotherapy_drug) %>% 
  # # mutate(n = n()) %>%
  # # mutate(tertile_pre_firstseqsample = ntile(cumulative_dose_pre_firstseqsample, 3))
  # mutate(tertile_pre_firstseqsample = case_when(
  #   n() > 2                                             ~ ntile(cumulative_dose_pre_firstseqsample, 3),
  #   n() == 1 &
  #     chemotherapy_drug == "abraxane"                   ~ 1,
  #   n() == 2 &
  #     chemotherapy_drug %in% 
  #     c("epirubicin", "fluorouracil")                   ~ 2
  # )) %>% 
  # mutate(tertile_pre_lastseqsample = case_when(
  #   n() > 2                                             ~ ntile(cumulative_dose_pre_lastseqsample, 3),
  #   n() == 1 &
  #     chemotherapy_drug == "abraxane"                   ~ 1,
  #   n() == 2 &
  #     chemotherapy_drug %in% 
  #     c("epirubicin", "fluorouracil")                   ~ 2
  # )) %>% 
  # # mutate(tertile_cum_drug_cat = case_when(
  # #   n() > 1 &
  # #     ntile(cumulative_dose, 3) == 1                    ~ "Low",
  # #   n() > 1 &
  # #     ntile(cumulative_dose, 3) == 2                    ~ "Medium",
  # #   n() > 1 &
  # #     ntile(cumulative_dose, 3) == 3                    ~ "High",
  # # )) %>% 
  # 
  # # calculate score - sum scores for each drug in a specific drug class
  # group_by(mrn, narrow_drug_class_cytotoxic_only) %>% 
  # # mutate(n = n()) %>%
  # mutate(drug_class_score_pre_firstseqsample = case_when(
  #   !is.na(tertile_pre_firstseqsample)                  ~ sum(tertile_pre_firstseqsample, na.rm = TRUE)
  # )) %>% 
  # mutate(drug_class_score_pre_lastseqsample = case_when(
  #   !is.na(tertile_pre_lastseqsample)                   ~ sum(tertile_pre_lastseqsample, na.rm = TRUE)
  # )) %>% 
  # distinct(mrn, narrow_drug_class_cytotoxic_only, .keep_all = TRUE) %>% 
  # # select(-chemotherapy_drug) %>% 
  # # group_by(narrow_drug_class_cytotoxic_only) %>%
  # # mutate(n = n()) %>% 
  # # Bolton summed score
  # group_by(mrn, narrow_drug_class_cytotoxic_only) %>% 
  # mutate(sum_bolton_drug_score_pre_firstseqsample_per_class_patient = case_when(
  #   !is.na(drug_class_score_pre_firstseqsample)         ~ sum(drug_class_score_pre_firstseqsample, na.rm = TRUE)
  # )) %>% 
  # mutate(sum_bolton_drug_score_pre_lastseqsample_per_class_patient = case_when(
  #   !is.na(drug_class_score_pre_lastseqsample)          ~ sum(drug_class_score_pre_lastseqsample, na.rm = TRUE)
  # )) %>% 
  # # Patient's
  # group_by(mrn) %>% 
  # mutate(sum_all_drug_score_pre_firstseqsample_per_patient = case_when(
  #   !is.na(drug_class_score_pre_firstseqsample)         ~ sum(drug_class_score_pre_firstseqsample, na.rm = TRUE)
  # )) %>% 
  # mutate(sum_all_drug_score_pre_lastseqsample_per_patient = case_when(
  #   !is.na(drug_class_score_pre_lastseqsample)          ~ sum(drug_class_score_pre_lastseqsample, na.rm = TRUE)
  # )) %>% 
  # ungroup() %>% 
  # # # use cut() instead of ntile() to attribute the ties in values
  # mutate(tertile_patient_pre_firstseqsample = cut(sum_all_drug_score_pre_firstseqsample_per_patient,
  #                            breaks = quantile(sum_all_drug_score_pre_firstseqsample_per_patient, 
  #                                              probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
  #                            include.lowest = TRUE,
  #                            labels = c("Low", "Medium", "High"))) %>%
  # mutate(tertile_patient_pre_lastseqsample = cut(sum_all_drug_score_pre_lastseqsample_per_patient,
  #                                               breaks = quantile(sum_all_drug_score_pre_lastseqsample_per_patient, 
  #                                                                 probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
  #                                               include.lowest = TRUE,
  #                                               labels = c("Low", "Medium", "High"))) %>% 
  full_join(sequenced_patient_data %>%
              select(mrn, deidentified_patient_id) %>%
              distinct() %>%
              mutate(mrn = as.character(mrn)),
            .,
            by = "mrn") #%>%
  # # select(mrn, deidentified_patient_id, narrow_drug_class_cytotoxic_only, 
  # #        sum_bolton_drug_score_pre_firstseqsample_per_class_patient, 
  # #        sum_bolton_drug_score_pre_lastseqsample_per_class_patient,
  # #        sum_all_drug_score_pre_firstseqsample_per_patient,
  # #        tertile_patient_pre_firstseqsample,
  # #        sum_all_drug_score_pre_lastseqsample_per_patient,
  # #        tertile_patient_pre_lastseqsample
  # #        ) %>% 
  # filter(!is.na(narrow_drug_class_cytotoxic_only))

# Make regimen
drug_summary <- drug_dose1 %>% 
  select(deidentified_patient_id, 
         chemotherapy_drug, narrow_drug_class_cytotoxic_only) %>% 
  group_by(deidentified_patient_id) %>% 
  summarise_at(vars(chemotherapy_drug, narrow_drug_class_cytotoxic_only), 
               str_c, collapse = "; ")
```

```{r}
drug_dose_dat <- data_merge %>% 
  left_join(., drug_summary, by = c("deidentified_patient_id")) 

drug_dose_dat %>% 
  select(tertile_patient_pre_firstseqsample, 
         chemotherapy_drug, 
         tertile_patient_pre_firstseqsample) %>% 
  tbl_summary(by = tertile_patient_pre_firstseqsample,
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels()

drug_dose_dat %>% 
  select(deidentified_patient_id, 
         tertile_patient_pre_firstseqsample, 
         chemotherapy_drug)

write_csv(drug_dose_dat %>% 
  select(deidentified_patient_id, 
         tertile_patient_pre_firstseqsample, 
         chemotherapy_drug), 
  paste0(here::here(),
          "/processed data/BreastCH_HomogenousPatientDrugsTertileGroup_20251121.csv"))
```

## 1. cyclophosphamide; docetaxel
```{r}
cyclo_doce <- drug_dose_dat %>% 
  filter(chemotherapy_drug == "cyclophosphamide; docetaxel")

tbl1 <- tbl_uvregression(
  cyclo_doce,
  method = coxph,
  y = Surv(time = pfs_time_months, event = pfs_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", "cond", 
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = cyclo_doce$pfs_time_months,
                   event = cyclo_doce$pfs_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + cond +
                tertile_patient_pre_firstseqsample,
              data = cyclo_doce)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

```{r, fig.height = 7}
ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ cond,
                   data=cyclo_doce),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```

```{r, fig.height = 7}
ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ cond,
                   data=cyclo_doce),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

## 2. cyclophosphamide; docetaxel & Low tertile 
```{r}
cyclo_doce <- drug_dose_dat %>% 
  filter(chemotherapy_drug == "cyclophosphamide; docetaxel" & 
           tertile_patient_pre_firstseqsample == "Low")

tbl1 <- tbl_uvregression(
  cyclo_doce,
  method = coxph,
  y = Surv(time = pfs_time_months, event = pfs_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", "cond"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = cyclo_doce$pfs_time_months,
                   event = cyclo_doce$pfs_event) ~ age_at_sample + tnm_stage_cat + 
                er + pr + her2 + cond,
              data = cyclo_doce)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2), 
          tab_spanner = c("**Univariable**", 
                          "**Multivariable**"))
```

```{r, fig.height = 7}
ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ cond,
                   data=cyclo_doce),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```

```{r, fig.height = 7}
ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ cond,
                   data=cyclo_doce),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>

## 3. cyclophosphamide; doxorubicin; paclitaxel
```{r}
cyclo_dox_pac <- drug_dose_dat %>%
  filter(chemotherapy_drug == "cyclophosphamide; doxorubicin; paclitaxel")

tbl1 <- tbl_uvregression(
  cyclo_dox_pac,
  method = coxph,
  y = Surv(time = pfs_time_months, event = pfs_event),
  exponentiate = TRUE,
  add_estimate_to_reference_rows = TRUE,
  include = c("age_at_sample", "tnm_stage_cat",
              "er", "pr", "her2", "cond",
              "tertile_patient_pre_firstseqsample"),
  # pvalue_fun = label_style_pvalue(digits = 2)
  ) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl2 <- coxph(Surv(time = cyclo_dox_pac$pfs_time_months,
                   event = cyclo_dox_pac$pfs_event) ~ age_at_sample + tnm_stage_cat +
                er + pr + her2 + #cond +
                tertile_patient_pre_firstseqsample,
              data = cyclo_dox_pac)  %>%
  tbl_regression(exponentiate = TRUE,
                 add_estimate_to_reference_rows = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()

tbl_merge(list(tbl1, tbl2),
          tab_spanner = c("**Univariable**",
                          "**Multivariable**"))
```


```{r, fig.height = 7}
ggsurvplot(survfit(Surv(os_time_from_tx_start_months, os_event) ~ cond,
                   data=cyclo_dox_pac),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```

```{r, fig.height = 7}
ggsurvplot(survfit(Surv(pfs_time_months, pfs_event) ~ cond,
                   data=cyclo_dox_pac),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) %++% guides(colour = guide_legend(ncol = 1))
```
<br>











