---
title: "Immunotherapy Samples"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
library(ggforce)
library(kableExtra)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r}
blood_patients <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/blood_patients.rds")
```

# I. Overall Data Selection
## Sample types selection
- collection_site_tissue_type == blood
- sample_type == Buffy Coat, Genomic DNA, MNC, MNC less CD138+, Unprocessed Liquid Tissue

## Patient characteristics
- primary_site_group == "Breast" (keep only breast diagnosis date)

# II. Repartition of Samples Before and After Immunotherapy within Time Limits

We didn't select the samples depending hormonetherapy.

Summary criteria for sample selection:  

* Immunotherapy:  
  + Pre-sample must be closest to Immunotherapy start from -365 days to Immunotherapy date 
     + also needs to be before chemo or no chemo was given, before radiation or no radiation was given   
  + Sequential sample must be after 365 days from Immunotherapy start  
     + also needs to be before chemo or no chemo was given, before radiation or no radiation was given  

We can change these criteria if we want to add more samples.  
```{r}
blood_patients <- blood_patients %>%
  mutate(blood_bf_immuno = case_when(
    specimen_collection_date <= (immunotherapy_start_date_1 - days(365))  ~ "too early",
    specimen_collection_date <= (immunotherapy_start_date_1 + days(5)) &
      (immunotherapy_start_date_1 <= chemotherapy_start_date_1 |
         is.na(chemotherapy_start_date_1)) &
      (specimen_collection_date <= radiation_start_date_1 |
      is.na(radiation_start_date_1))                                     ~ "Yes",
    specimen_collection_date > immunotherapy_start_date_1                 ~ "No",
    is.na(immunotherapy_start_date_1)                                     ~ "not administred",
    TRUE                                                                 ~ "No"
  )) %>%
  mutate(across(contains("blood_bf_"), ~ factor(., levels = c("Yes", "No", "not administred")))) %>%

  arrange(mrn, deidentified_patient_id, specimen_collection_date) %>%
  select(mrn, deidentified_patient_id, specimen_collection_date,
         "blood_bf_immuno", immunotherapy_start_date_1, chemotherapy_start_date_1, radiation_start_date_1,
         everything())
```


```{r code pre and sequential}
blood_patients2 <- blood_patients %>% 
  ###### Chemo samples ----
  mutate(had_good_sample_immuno = case_when(
    blood_bf_immuno == "Yes"                    ~ "Yes"
  )) %>%
  group_by(deidentified_patient_id, had_good_sample_immuno) %>%
  mutate(closest_immuno_presample_date = case_when(
    had_good_sample_immuno == "Yes"            ~ last(specimen_collection_date)
  )) 

blood_patients3 <- blood_patients2 %>%
  mutate(interval_presample_immuno = case_when(
    closest_immuno_presample_date == specimen_collection_date  ~ 
      (interval(start = closest_immuno_presample_date, end = immunotherapy_start_date_1) /
                 duration(n = 1, units = "days"))
  )) %>%
  mutate(presample_id_immuno = case_when(
    !is.na(interval_presample_immuno)       ~ sample_id
  )) %>% 
  mutate(presample_date_immuno = case_when(
    !is.na(presample_id_immuno)                                     ~ specimen_collection_date
  )) 

blood_patients4 <- blood_patients3 %>% 
  
  # Find sequential samples
  mutate(presample_immuno_within_limits = had_good_sample_immuno) %>% 
  group_by(deidentified_patient_id) %>%
  fill(presample_immuno_within_limits, #interval_presample_immuno,
       closest_immuno_presample_date, .direction = "updown") %>%

  mutate(seq_sample_immuno = case_when(
    presample_immuno_within_limits == "Yes" &
      blood_bf_immuno == "No" &
      specimen_collection_date >= (immunotherapy_start_date_1 + days(365)) &
      (specimen_collection_date <= chemotherapy_start_date_1 |
         is.na(chemotherapy_start_date_1)) &
      (specimen_collection_date <= radiation_start_date_1 |
      is.na(radiation_start_date_1))                                ~ "Yes",
    TRUE                                                            ~ "No"
  )) %>%
  group_by(mrn, seq_sample_immuno) %>% 
  mutate(is_first_seq_sample = case_when(
    seq_sample_immuno == "Yes" &
    specimen_collection_date == first(specimen_collection_date)     ~ "Yes",
    TRUE                                                            ~ "No"
  )) %>% 
  ungroup() %>% 
  
  mutate(seqsample_id_immuno = case_when(
    is_first_seq_sample == "Yes"                                    ~ sample_id
  )) %>% 
  
  mutate(seqsample_date_immuno = case_when(
    !is.na(seqsample_id_immuno)                                     ~ specimen_collection_date
  )) %>% 
  mutate(seqsample_before_2nd_regimen = case_when(
    !is.na(seqsample_id_immuno) &
      specimen_collection_date <= immunotherapy_start_date_2        ~ "sample before 2nd regimen",
    !is.na(seqsample_id_immuno) &
      specimen_collection_date > immunotherapy_start_date_2         ~ "sample after"
  )) %>% 
  mutate(have_good_seqsample_immuno = case_when(
    seq_sample_immuno == "Yes"                                       ~ "Yes"
  )) %>% 
  fill(have_good_seqsample_immuno, .direction = "updown") %>% 
  mutate(time_immuno_seqsample = case_when(
    seq_sample_immuno == "Yes"        ~ (
      interval(start = immunotherapy_start_date_1, end = specimen_collection_date) /
        duration(n= 1, units = "days")
    ))) %>%
  mutate(interval_prepost_sample_immuno = case_when(
    !is.na(seqsample_id_immuno)                                     ~
      abs(interval(start = closest_immuno_presample_date, end = specimen_collection_date) /
                 duration(n = 1, units = "days"))
  )) 

blood_patients2 <- blood_patients4 %>%
  mutate(across(where(is.numeric), ~ round(., 2)))

```

# III. Patients with Blood before and after radiotharapy
## 1. Patients table 
```{r}
tbl1_1 <-
  blood_patients2 %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(Immunotherapy = presample_immuno_within_limits) %>%
  tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>%
  modify_header(update = list(
  stat_0 ~ '**Patients with Blood before Immunotherapy**'
))
tbl1_2 <- blood_patients2 %>%
#   filter(!is.na(seqsample_id_immuno)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(Immunotherapy = seq_sample_immuno) %>%
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))%>%
  modify_header(update = list(
  stat_0 ~ '**Patients with Sequential Blood after Immunotherapy**'
))

tbl_merge(list(tbl1_1, tbl1_2)) %>%
   modify_header(list(label ~ "**Treatment Type**")) %>%
  modify_spanning_header(everything() ~ NA_character_)

```
