---
title: "Patients with multiple samples"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(tidyverse)
library(lubridate)
library(gtsummary)
library(viridis)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r}
# blood_patients <- 
#   read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/blood_patients.rds")
blood_patients2 <- read_rds(paste0(here::here(), "/blood_patients2.rds"))
```

# All breast cancers with blood sample
```{r}
blood_patients2 %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = date_of_first_sample) %>% 
  ggplot(aes(x =time_from_first_sample_to_any_next_sample))+
  geom_histogram(binwidth = 1, fill= "darkblue") +
  # xlim(0, 400)+
  labs(x="Time from First Sample to \nAny Following Sample (in years)", 
       y="Number of Patient",
       caption = "Each bar represents 1 year")
```


```{r, fig.height= 6}
blood_patients2 %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = date_of_first_sample) %>% 
  
  ggplot(aes(x = time_from_first_sample_to_any_next_sample, fill = as.factor(sample_seq_number)))+
  geom_histogram(binwidth = 1, position = "stack") +
  # xlim(0, 400)+
  scale_fill_viridis(discrete=T, name = "Time between the 1st and ... sample")+
  labs(x="Time from First Sample to \nAny Following Sample (in years)", 
       y="Number of Patient",
       caption = "Each bar represents 1 year")+
  theme(legend.position = "bottom")

blood_patients2 %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = date_of_first_sample) %>% 
  
  ggplot(aes(x = time_from_first_sample_to_any_next_sample, fill = as.factor(sample_seq_number)))+
  geom_histogram(binwidth = 1, position = "stack") +
  coord_cartesian(ylim = c(0, 40))+
  scale_fill_viridis(discrete=T, name = "Time between the 1st and ... sample")+
  labs(x="Time from First Sample to \nAny Following Sample (in years)", 
       y="Number of Patient",
       caption = "Each bar represents 1 year")+
  theme(legend.position = "bottom")
```


```{r}
a <- blood_patients2 %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "days"), .after = date_of_first_sample)
a %>%   
  select(time_from_first_sample_to_any_next_sample, sample_seq_number) %>% 
  tbl_summary(statistic =list(all_continuous() ~ "{mean} ({min}, {max})")) %>% 
  modify_caption("**Patients** (N = {a %>% distinct(deidentified_patient_id) %>% nrow()})")
```

# With subsequent myeloid malignancies
```{r list of subsequent cancer}
# b <- breast_info %>%
#   filter(primary_site_group_desc != "BREAST") %>%
#   select(patient_id, histology_cd, histology_desc, primary_site_cd, primary_site_group_desc) %>%
#   arrange(primary_site_group_desc, histology_desc) %>% 
#   left_join(., Demographic %>% select(patient_id, mrn))
# write_csv(b, "patient subsequent cancer after breast cancer - hossein.csv")
subsequent_cancer <- 
  read_csv("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/patient subsequent cancer after breast cancer - hossein.csv")
subsequent_cancer_of_interest <- 
  readxl::read_xlsx("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/Subsequent cancers after breast_02.05.24.xlsx",
                    sheet = "Myeloid malignancies", col_names = FALSE) %>% 
  `colnames<-`("histology_desc")
subsequent_cancer <- subsequent_cancer %>% 
  inner_join(., subsequent_cancer_of_interest,
             by = "histology_desc") %>% 
  select(mrn, histology_desc, dx_dt) %>% 
  mutate(mrn = as.character(mrn)) %>% 
  group_by(mrn) %>% 
  summarise_at(vars(histology_desc, dx_dt), 
               str_c, collapse = "; ") %>% 
  ungroup() %>% 
  mutate(dx_count = sapply(strsplit(dx_dt, "; "), length)) %>% 
  
  separate_wider_delim(dx_dt, 
                       delim = "; ", 
                       names = c(paste("date_of_myeloid", 1:max(.$dx_count), sep = "")), 
                       too_few = "align_start", too_many = "drop", 
                       cols_remove = TRUE) %>% 
  separate_wider_delim(histology_desc, 
                       delim = "; ", 
                       names = c(paste("histology_desc", 1:max(.$dx_count), sep = "")), 
                       too_few = "align_start", too_many = "drop", 
                       cols_remove = TRUE)

blood_patients2_subset <- blood_patients2 %>% 
  filter(mrn %in% c(subsequent_cancer$mrn)) %>% 
  inner_join(subsequent_cancer, .,
             by = "mrn")
```

None of these patients have sequential samples which would have fit the first criteria we had. So it is highly possible they were never sequenced.
```{r}
blood_patients2_subset %>% 
  select(mrn, had_chemo, had_rad, had_hormone, had_immuno,
         presample_id_chemo, seqsample_id_chemo,
         presample_id_rad, seqsample_id_rad,
         presample_id_chemorad,
         seqsample_id_Achemo,
         seqsample_id_Achemorad,
         presample_id_hormone,
         seqsample_id_hormone) %>% 
  arrange(had_chemo, presample_id_chemo)
```

```{r}
blood_patients2_subset %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = date_of_first_sample) %>% 
  ggplot(aes(x =time_from_first_sample_to_any_next_sample))+
  geom_histogram(binwidth = 1, fill= "darkblue") +
  # xlim(0, 400)+
  labs(x="Time from First Sample to \nAny Following Sample (in years)", 
       y="Number of Patient",
       caption = "Each bar represents 1 year")
```


```{r, fig.height= 6}
blood_patients2_subset %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = date_of_first_sample) %>% 
  
  ggplot(aes(x = time_from_first_sample_to_any_next_sample, fill = as.factor(sample_seq_number)))+
  geom_histogram(binwidth = 1, position = "stack") +
  # xlim(0, 400)+
  scale_fill_viridis(discrete=T, name = "Time between the 1st and ... sample")+
  labs(x="Time from First Sample to \nAny Following Sample (in years)", 
       y="Number of Patient",
       caption = "Each bar represents 1 year")+
  theme(legend.position = "bottom")
```


```{r}
a <- blood_patients2_subset %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  filter(sample_seq_number != 1) %>% 
  mutate(time_from_first_sample_to_any_next_sample = interval(start = date_of_first_sample, end = specimen_collection_date) /
           duration(n = 1, unit = "days"), .after = date_of_first_sample)
a %>%   
  select(time_from_first_sample_to_any_next_sample, sample_seq_number) %>% 
  tbl_summary(statistic =list(all_continuous() ~ "{mean} ({min}, {max})")) %>% 
  modify_caption("**Patients** (N = {a %>% distinct(deidentified_patient_id) %>% nrow()})")
```

Time from samples to myeloid malignancy
```{r}
blood_patients2_subset %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date) %>% 
  # mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  # filter(sample_seq_number == 1) %>% 
  
  mutate(time_from_malignancy_to_sample_myeloid = interval(start = date_of_myeloid1, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = sample_seq_number) %>% 
  
  ggplot(aes(x = time_from_malignancy_to_sample_myeloid, fill = as.factor(sample_seq_number)))+
  geom_histogram(binwidth = 1, position = "stack") +
  # xlim(0, 400)+
  scale_fill_viridis(discrete=T, name = "Time between the 1st and ... sample")+
  labs(x="Time from First Sample to \nAny Following Sample (in years)", 
       y="Number of Patient",
       caption = "Each bar represents 1 year")+
  theme(legend.position = "bottom")
```


```{r, fig.height=11, fig.width=12}
blood_patients2_subset %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date,
         sample_seq_number = factor(sample_seq_number, levels = c("3", "2", "1"))) %>% 
  # mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  # filter(sample_seq_number == 1) %>% 
  
  mutate(time_from_malignancy_to_sample_myeloid = interval(start = date_of_myeloid1, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = sample_seq_number) %>% 
  mutate(text = paste("Patient: ", mrn, "\n", sample_seq_number, " sample ", 
                      "collected on ", as.Date(specimen_collection_date), 
                      "\n",round(time_from_malignancy_to_sample_myeloid, 2), " from myeloid malingnancy", sep="")) %>% 

  ggplot(aes(x = time_from_malignancy_to_sample_myeloid, y = mrn, color = sample_seq_number, group = mrn))+
  geom_point(shape = 2, position=position_jitter(width = 0.2, height = 0))+
  geom_line() +
  # xlim(0, 400)+
  geom_vline(xintercept = 0, color = "grey")+
  labs(x="Time from Sample to \nMyeloid Malingnancy (in years)", 
       y="Patient")+
  theme(legend.position = "bottom", text = element_text(size = 17))
```


```{r, fig.height=9, fig.width=9}
library(plotly)
plot <- blood_patients2_subset %>% 
  arrange(deidentified_patient_id, specimen_collection_date) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_seq_number = row_number(), .after = specimen_collection_date,
         sample_seq_number = factor(sample_seq_number, levels = c("3", "2", "1"))) %>% 
  # mutate(date_of_first_sample = first(specimen_collection_date), .after = sample_seq_number) %>% 
  ungroup() %>% 
  mutate(tt = case_when(
    sample_seq_number == 1    ~ "st",
    sample_seq_number == 2    ~ "nd",
    sample_seq_number == 3    ~ "rd"
  )) %>% 
  # filter(sample_seq_number == 1) %>% 
  
  mutate(time_from_malignancy_to_sample_myeloid = interval(start = date_of_myeloid1, end = specimen_collection_date) /
           duration(n = 1, unit = "years"), .after = sample_seq_number) %>% 
  mutate(sequence = case_when(
    time_from_malignancy_to_sample_myeloid <= 0    ~ "before",
    time_from_malignancy_to_sample_myeloid > 0     ~ "after"
  )) %>% 
  mutate(text = paste("Patient:", mrn, "\n", sample_seq_number, tt, "sample", 
                      "collected on", as.Date(specimen_collection_date), 
                      "\n",round(time_from_malignancy_to_sample_myeloid, 2), "years",
                      sequence, "myeloid malingnancy", sep=" ")) %>% 

  ggplot(aes(x = time_from_malignancy_to_sample_myeloid, y = mrn, color = sample_seq_number, text = text))+
  geom_point(shape = 2, position=position_jitter(width = 0.2, height = 0))+
  geom_line() +
  geom_vline(xintercept = 0, color = "grey")+
  # scale_fill_viridis(discrete=T, name = "Time between myeloid malignancy and samples")+
  labs(x="Time from Sample to \nMyeloid Malingnancy (in years)", 
       y="Patient")+
  guides(colour = "none")

plot <- ggplotly(plot, tooltip="text")
plot
```









