---
title: "Radiation Samples"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
library(ggforce)
library(kableExtra)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r}
blood_patients <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/blood_patients.rds")
```

# I. Overall Data Selection
## Sample types selection
- collection_site_tissue_type == blood
- sample_type == Buffy Coat, Genomic DNA, MNC, MNC less CD138+, Unprocessed Liquid Tissue

## Patient characteristics
- primary_site_group == "Breast" (keep only breast diagnosis date)

# II. Repartition of Samples Before and After Radiotherapy within Time Limits

In the previous sample selection, we focused on the chemotherapy and radiation treatment so we didn't select the samples depending on immunotherapy or hormonetherapy. I did the same here.  

Summary criteria for sample selection:  

* Radiation only:  
  + Pre-sample must be closest to radiation start from -365 days to radiation date 
     + also needs to be before chemo or no chemo was given  
  + Sequential sample must be within a year but 100 days after radiation  
     + also needs to be before chemo or no chemo was given  
We can change these criteria if we want to add more samples.  
```{r}
# blood_patients <- blood_patients %>% 
#   mutate(blood_bf_chemo = case_when(
#     specimen_collection_date <= (chemotherapy_start_date_1 - days(365))  ~ "too early",
#     specimen_collection_date <= (chemotherapy_start_date_1 + days(5)) &
#       # specimen_collection_date <= specimen_collection_date - 365
#       (specimen_collection_date <= radiation_start_date_1 |
#       is.na(radiation_start_date_1))                                     ~ "Yes",
#     specimen_collection_date > chemotherapy_start_date_1                 ~ "No",
#     is.na(chemotherapy_start_date_1)                                     ~ "not administred",
#     TRUE                                                                 ~ "No"
#   )) %>% 
#   mutate(blood_bf_hormone = case_when(
#     specimen_collection_date <= (hormone_therapy_start_date_1 - days(365))  ~ "too early",
#     specimen_collection_date <= (hormone_therapy_start_date_1 + days(5)) &
#     (specimen_collection_date <= chemotherapy_start_date_1 |
#       is.na(chemotherapy_start_date_1)) &
#       (specimen_collection_date <= radiation_start_date_1 |
#       is.na(radiation_start_date_1))                                        ~ "Yes",
#     specimen_collection_date > hormone_therapy_start_date_1                 ~ "No",
#     is.na(hormone_therapy_start_date_1)                                     ~ "not administred",
#     TRUE                                                                    ~ "No"
#   )) %>%
#   mutate(blood_bf_rad = case_when(
#     specimen_collection_date <= (radiation_start_date_1 - days(365))    ~ "too early",
#     specimen_collection_date <= radiation_start_date_1 &
#       (specimen_collection_date <= chemotherapy_start_date_1 |
#       is.na(chemotherapy_start_date_1))                                 ~ "Yes",
#     specimen_collection_date > radiation_start_date_1                   ~ "No",
#     is.na(radiation_start_date_1)                                       ~ "not administred",
#     TRUE                                                                ~ "No"
#   )) %>% 
#   mutate(blood_bf_chemo_rad = case_when(
#     specimen_collection_date <= (chemotherapy_start_date_1 - days(365))             ~ "too early",
#     specimen_collection_date <= (chemotherapy_start_date_1 + days(5)) &
#       specimen_collection_date <= (radiation_start_date_1 + days(5))                ~ "Yes",
#     specimen_collection_date > (chemotherapy_start_date_1 + days(5)) |
#       specimen_collection_date > (radiation_start_date_1 + days(5))                 ~ "No",
#     is.na(chemotherapy_start_date_1) |
#       is.na(radiation_start_date_1)                                     ~ "both not administred",
#     TRUE                                                                ~ "No"
#   )) %>% 
#   mutate(first_chemo_rad_date = case_when(
#     chemotherapy_start_date_1 <= radiation_start_date_1                 ~ chemotherapy_start_date_1,
#     chemotherapy_start_date_1 > radiation_start_date_1                  ~ radiation_start_date_1,
#     TRUE                                                                ~ NA_Date_
#   )) %>% 
#   
#   mutate(across(contains("blood_bf_"), ~ factor(., levels = c("Yes", "No", "not administred")))) %>% 
#  
#   arrange(mrn, deidentified_patient_id, specimen_collection_date) %>% 
#   select(mrn, deidentified_patient_id, specimen_collection_date, 
#          "blood_bf_chemo", "blood_bf_rad",
#          "blood_bf_hormone", 
#          "blood_bf_chemo_rad",
#          everything())
# ```
# 
# 
# ```{r code pre and sequential}
# # blood_patients2 <- blood_patients %>% 
# #   ###### Chemo samples ----
# #   mutate(had_good_sample_chemo = case_when(
# #     blood_bf_chemo == "Yes"                    ~ "Yes"
# #   )) %>%
# #   group_by(deidentified_patient_id, had_good_sample_chemo) %>%
# #   mutate(closest_chemo_presample_date = case_when(
# #     had_good_sample_chemo == "Yes"            ~ last(specimen_collection_date)
# #   )) 
# # 
# # blood_patients3 <- blood_patients2 %>%
# #   mutate(interval_presample_chemo = case_when(
# #     closest_chemo_presample_date == specimen_collection_date  ~ 
# #       (interval(start = closest_chemo_presample_date, end = chemotherapy_start_date_1) /
# #                  duration(n = 1, units = "days"))
# #   )) %>%
# #   mutate(presample_id_chemo = case_when(
# #     !is.na(interval_presample_chemo)       ~ sample_id
# #   )) %>% 
# #   mutate(presample_date_chemo = case_when(
# #     !is.na(presample_id_chemo)                                     ~ specimen_collection_date
# #   )) 
# # 
# # blood_patients4 <- blood_patients3 %>% 
# #   
# #   # Find sequential samples
# #   mutate(presample_chemo_within_limits = had_good_sample_chemo) %>% 
# #   group_by(deidentified_patient_id) %>%
# #   fill(presample_chemo_within_limits, #interval_presample_chemo,
# #        closest_chemo_presample_date, .direction = "updown") %>%
# #   mutate(seq_sample_chemo = case_when(
# #     presample_chemo_within_limits == "Yes" &
# #       blood_bf_chemo == "No" &
# #       specimen_collection_date >= chemotherapy_end_date1_1 &
# #       specimen_collection_date <= (chemotherapy_start_date_1 + days(365)) &
# #       (specimen_collection_date <= radiation_start_date_1 |
# #       is.na(radiation_start_date_1))                                ~ "Yes",
# #     TRUE                                                            ~ "No"
# #   )) %>%
# #   group_by(mrn, seq_sample_chemo) %>% 
# #   mutate(is_first_seq_sample = case_when(
# #     seq_sample_chemo == "Yes" &
# #     specimen_collection_date == first(specimen_collection_date)     ~ "Yes",
# #     TRUE                                                            ~ "No"
# #   )) %>% 
# #   ungroup() %>% 
# #   
# #   mutate(seqsample_id_chemo = case_when(
# #     is_first_seq_sample == "Yes"                                    ~ sample_id
# #   )) %>% 
# #   
# #   mutate(seqsample_date_chemo = case_when(
# #     !is.na(seqsample_id_chemo)                                     ~ specimen_collection_date
# #   )) %>% 
# #   mutate(seqsample_before_2nd_regimen = case_when(
# #     !is.na(seqsample_id_chemo) &
# #       specimen_collection_date <= chemotherapy_start_date_2        ~ "sample before 2nd regimen",
# #     !is.na(seqsample_id_chemo) &
# #       specimen_collection_date > chemotherapy_start_date_2         ~ "sample after"
# #   )) %>% 
# #   mutate(have_good_seqsample_chemo = case_when(
# #     seq_sample_chemo == "Yes"                                       ~ "Yes"
# #   )) %>% 
# #   fill(have_good_seqsample_chemo, .direction = "updown") %>% 
# #   mutate(time_chemo_seqsample = case_when(
# #     seq_sample_chemo == "Yes"        ~ (
# #       interval(start = chemotherapy_start_date_1, end = specimen_collection_date) /
# #         duration(n= 1, units = "days")
# #     ))) %>%
# #   mutate(interval_prepost_sample_chemo = case_when(
# #     !is.na(seqsample_id_chemo)                                     ~
# #       abs(interval(start = closest_chemo_presample_date, end = specimen_collection_date) /
# #                  duration(n = 1, units = "days"))
# #   )) 
# 
# blood_patients5 <- blood_patients %>%
#   
#   ###### Rad samples ----
#   mutate(had_good_sample_rad = case_when(
#     blood_bf_rad == "Yes"              ~ "Yes"
#   )) %>%
#   group_by(deidentified_patient_id, had_good_sample_rad) %>%
#   mutate(closest_rad_presample_date = case_when(
#     had_good_sample_rad == "Yes"       ~ last(specimen_collection_date)
#   )) %>%
#   mutate(interval_presample_rad = case_when(
#     closest_rad_presample_date == specimen_collection_date  ~ 
#       (interval(start = closest_rad_presample_date, end = radiation_start_date_1) /
#                  duration(n = 1, units = "days"))
#   )) %>%
#   mutate(presample_id_rad = case_when(
#     !is.na(interval_presample_rad)                           ~ sample_id
#   )) %>% 
#   mutate(presample_date_rad = case_when(
#     !is.na(presample_id_rad)                                        ~ specimen_collection_date
#   )) 
# 
# blood_patients4 <- blood_patients5 %>%
#   
#   # Find sequential samples
#   mutate(presample_rad_within_limits = had_good_sample_rad) %>%
#   group_by(deidentified_patient_id) %>%
#   fill(presample_rad_within_limits, #interval_presample_rad,
#        closest_rad_presample_date, .direction = "updown") %>%
#   mutate(seq_sample_rad = case_when(
#     presample_rad_within_limits == "Yes" &
#       blood_bf_rad == "No" &
#       specimen_collection_date > (radiation_start_date_1 + days(100)) &
#       specimen_collection_date <= (radiation_start_date_1 + days(365)) &
#       (specimen_collection_date <= chemotherapy_start_date_1 |
#       is.na(chemotherapy_start_date_1))                              ~ "Yes",
#     TRUE ~ "No"
#   )) 
# 
# blood_patients3 <- blood_patients4 %>%
#   
#   group_by(mrn, seq_sample_rad) %>% 
#   mutate(is_first_seq_sample = case_when(
#     seq_sample_rad == "Yes" &
#     specimen_collection_date == first(specimen_collection_date)     ~ "Yes",
#     TRUE                                                            ~ "No"
#   )) %>% 
#   ungroup() 
# 
# blood_patients2 <- blood_patients3 %>%
#   
#   mutate(seqsample_id_rad = case_when(
#     is_first_seq_sample == "Yes"                                    ~ sample_id
#   )) %>% 
#   mutate(seqsample_date_rad = case_when(
#     !is.na(seqsample_id_rad)                                        ~ specimen_collection_date
#   )) %>% 
#   mutate(have_good_seqsample_rad = case_when(
#     seq_sample_rad == "Yes"                                         ~ "Yes"
#   )) %>% 
#   fill(have_good_seqsample_rad, .direction = "updown") %>% 
#   mutate(time_rad_seqsample = case_when(
#     seq_sample_rad == "Yes"        ~ (
#       interval(start = radiation_start_date_1, end = specimen_collection_date) /
#         duration(n= 1, units = "days")
#   ))) %>%
#   mutate(interval_prepost_sample_rad =
#            abs(interval(start = closest_rad_presample_date, end = specimen_collection_date) /
#                  duration(n = 1, units = "days"))) 
# 
# # blood_patients3 <- blood_patients4 %>%
# #   ###### hormone samples ----
# #   mutate(had_good_sample_hormone = case_when(
# #     blood_bf_hormone == "Yes"              ~ "Yes"
# #   )) %>%
# #   group_by(deidentified_patient_id, had_good_sample_hormone) %>%
# #   mutate(closest_hormone_presample_date = case_when(
# #     had_good_sample_hormone == "Yes"       ~ last(specimen_collection_date)
# #   )) %>%
# #   mutate(interval_presample_hormone = case_when(
# #     closest_hormone_presample_date == specimen_collection_date  ~ 
# #       (interval(start = closest_hormone_presample_date, end = hormone_therapy_start_date_1) /
# #                  duration(n = 1, units = "days"))
# #   )) %>%
# #   mutate(presample_id_hormone = case_when(
# #     !is.na(interval_presample_hormone)                           ~ sample_id
# #   )) %>% 
# #   mutate(presample_date_hormone = case_when(
# #     !is.na(presample_id_hormone)                                     ~ specimen_collection_date
# #   )) %>% 
# #   
# #   # Find sequential samples
# #   mutate(presample_hormone_within_limits = had_good_sample_hormone) %>%
# #   group_by(deidentified_patient_id) %>%
# #   fill(presample_hormone_within_limits, #interval_presample_hormone,
# #        closest_hormone_presample_date, .direction = "updown") %>%
# #   mutate(seq_sample_hormone = case_when(
# #     presample_hormone_within_limits == "Yes" &
# #       blood_bf_hormone == "No" &
# #       specimen_collection_date >= (hormone_therapy_start_date_1 + days(85)
# #                                   ) &
# #       specimen_collection_date <= (hormone_therapy_start_date_1 + days(365)
# #                                    ) &
# #       (specimen_collection_date <= chemotherapy_start_date_1 |
# #       is.na(chemotherapy_start_date_1)) &
# #       (specimen_collection_date <= radiation_start_date_1 |
# #       is.na(radiation_start_date_1))                                     ~ "Yes",
# #     TRUE ~ "No"
# #   )) %>%
# #   
# #   group_by(mrn, seq_sample_hormone) %>% 
# #   mutate(is_first_seq_sample = case_when(
# #     seq_sample_hormone == "Yes" &
# #     specimen_collection_date == first(specimen_collection_date)     ~ "Yes",
# #     TRUE                                                            ~ "No"
# #   )) %>% 
# #   ungroup() %>% 
# #   
# #   mutate(seqsample_id_hormone = case_when(
# #     is_first_seq_sample == "Yes"                                    ~ sample_id
# #   )) %>% 
# #   
# #   mutate(seqsample_date_hormone = case_when(
# #     !is.na(seqsample_id_hormone)                                      ~ specimen_collection_date
# #   )) %>% 
# #   mutate(seqsample_before_2nd_regimen_hormone = case_when(
# #     !is.na(seqsample_id_hormone) &
# #       specimen_collection_date <= hormone_therapy_start_date_2        ~ "sample before 2nd regimen",
# #     !is.na(seqsample_id_hormone) &
# #       specimen_collection_date > hormone_therapy_start_date_2         ~ "sample after"
# #   )) %>% 
# #   mutate(have_good_seqsample_hormone = case_when(
# #     seq_sample_hormone == "Yes"                                       ~ "Yes"
# #   )) %>% 
# #   fill(have_good_seqsample_hormone, .direction = "updown") %>% 
# #   mutate(time_hormone_seqsample = case_when(
# #     seq_sample_hormone == "Yes"        ~ (
# #       interval(start = hormone_therapy_start_date_1, end = specimen_collection_date) /
# #         duration(n= 1, units = "days")
# #   ))) %>%
# #   mutate(interval_prepost_sample_hormone =
# #            abs(interval(start = closest_hormone_presample_date, end = specimen_collection_date) /
# #                  duration(n = 1, units = "days"))) 
# 
# # blood_patients3 <- blood_patients4 %>%
# #   ###### C + Rad samples ---- 
# #   mutate(had_good_sample_chemrad = case_when(
# #     blood_bf_chemo_rad == "Yes"                              ~ "Yes"
# #   )) %>%
# #   group_by(deidentified_patient_id, had_good_sample_chemrad) %>%
# #   mutate(closest_chemorad_presample_date = case_when(
# #     had_good_sample_chemrad == "Yes"                         ~ last(specimen_collection_date)
# #   )) %>%
# #   mutate(interval_presample_chemrad = case_when(
# #     closest_chemorad_presample_date == specimen_collection_date  ~ 
# #       (interval(start = closest_chemorad_presample_date, end = first_chemo_rad_date) /
# #                  duration(n = 1, units = "days"))
# #   )) %>%
# #   mutate(presample_id_chemorad = case_when(
# #     !is.na(interval_presample_chemrad)                       ~ sample_id
# #   )) %>% 
# #   mutate(presample_date_chemorad = case_when(
# #     !is.na(presample_id_chemorad)                                     ~ specimen_collection_date
# #   )) %>% 
# #   
# #   # Find sequential samples
# #   mutate(presample_chemorad_within_limits = had_good_sample_chemrad) %>%
# #   group_by(deidentified_patient_id) %>%
# #   fill(presample_chemorad_within_limits, #interval_presample_chemrad,
# #        closest_chemorad_presample_date, .direction = "updown") %>%
# #   ungroup() %>% 
# #   # Intermediate var for seq sample
# #   mutate(interval_chemo_rad =
# #            abs(interval(start = chemotherapy_start_date_1, end = radiation_start_date_1) /
# #         duration(n= 1, units = "days"))
# #         ) %>%
# #   mutate(second_chemorad_date = first_chemo_rad_date + abs(interval_chemo_rad)) %>%
# #   mutate(seq_sample_chemorad = case_when(
# #     presample_chemorad_within_limits == "Yes" &
# #       blood_bf_chemo_rad == "No" &
# #       specimen_collection_date > (second_chemorad_date + days(100)) &
# #       specimen_collection_date <= (second_chemorad_date + days(365)) ~ "Yes",
# #     TRUE                                                            ~ "No"
# #   )) %>%
# #   
# #   group_by(mrn, seq_sample_chemorad) %>% 
# #   mutate(is_first_seq_sample = case_when(
# #     seq_sample_chemorad == "Yes" &
# #     specimen_collection_date == first(specimen_collection_date)     ~ "Yes",
# #     TRUE                                                            ~ "No"
# #   )) %>% 
# #   ungroup() %>% 
# #   
# #   mutate(seqsample_id_chemorad = case_when(
# #     is_first_seq_sample == "Yes"                                    ~ sample_id
# #   )) %>% 
# #   
# #   mutate(seqsample_date_chemorad = case_when(
# #     !is.na(seqsample_id_chemorad)                                   ~ specimen_collection_date
# #   )) %>% 
# #   mutate(have_good_seqsample_chemorad = case_when(
# #     seq_sample_chemorad == "Yes"                                    ~ "Yes"
# #   )) %>% 
# #   fill(have_good_seqsample_chemorad, .direction = "updown") %>% 
# #   # select(deidentified_patient_id, specimen_collection_date, had_good_sample_chemrad, closest_chemorad_presample_date, first_chemo_rad_date,second_chemrad_date,second_chemrad_date, seq_sample_chemorad, chemotherapy_start_date_1, radiation_start_date_1)
# #   mutate(time_chemorad_seqsample = case_when(
# #     seq_sample_chemorad == "Yes"                             ~ (
# #       interval(start = second_chemorad_date, end = specimen_collection_date) /
# #         duration(n= 1, units = "days")
# #   ))) %>%
# #   mutate(interval_prepost_sample_chemorad =
# #            abs(interval(start = closest_chemorad_presample_date, end = specimen_collection_date) /
# #                  duration(n = 1, units = "days"))) %>%
# 
#   ###### C then Rad sample ----
#   # mutate(had_good_sample_chemo_rad = case_when(
#   #   blood_bf_chemo_rad == "Yes" &
#   #     chemotherapy_start_date_1 < radiation_start_date_1     ~ "Yes"
#   # )) %>% 
#   # select(deidentified_patient_id, sample_id, specimen_collection_date, blood_bf_chemo_rad,
#   #        chemotherapy_start_date_1, radiation_start_date_1, had_good_sample_chemo_rad, 
#   #        first_chemo_rad_date, everything()) %>% 
#   # 
#   # 
#   # group_by(deidentified_patient_id, had_good_sample_chemo_rad) %>% 
#   # mutate(closest_chemo_rad_presample_date = case_when(
#   #   had_good_sample_chemo_rad == "Yes"                       ~ last(specimen_collection_date)
#   # )) %>% 
#   # mutate(interval_presample_chemo_rad = case_when(
#   #   closest_chemo_rad_presample_date == specimen_collection_date  ~ 
#   #     (interval(start = closest_chemo_rad_presample_date, end = chemotherapy_start_date_1) /
#   #                duration(n = 1, units = "days"))
#   # )) %>%
#   # mutate(presample_id_chemo_rad = case_when(
#   #   !is.na(interval_presample_chemo_rad)                     ~ sample_id
#   # )) %>% 
#   # mutate(presample_date_chemo_rad = case_when(
#   #   !is.na(presample_id_chemo_rad)                                     ~ specimen_collection_date
#   # )) %>% 
#   # 
#   # # Find seq sample after C and before R
#   # mutate(presample_chemo_rad_within_limits = had_good_sample_chemo_rad) %>%
#   # group_by(deidentified_patient_id) %>% 
#   # fill(presample_chemo_rad_within_limits, #closest_chemo_rad_presample_date, 
#   #      .direction = "updown") %>% 
#   # ungroup() %>% 
#   # mutate(seq_sample_Achemo_Brad = case_when(
#   #   presample_chemo_rad_within_limits == "Yes" &
#   #     blood_bf_chemo_rad == "No" &
#   #     # specimen_collection_date > (chemotherapy_start_date_1 + days(100)) &
#   #     # specimen_collection_date > chemotherapy_end_date1_1 &
#   #     specimen_collection_date > chemotherapy_start_date_1 &
#   #     specimen_collection_date <= radiation_start_date_1            ~ "Yes",
#   #   TRUE                                                            ~ "No"
#   # )) %>%
#   # group_by(deidentified_patient_id, seq_sample_Achemo_Brad) %>% 
#   # mutate(seqsample_id_Achemo = case_when(
#   #   seq_sample_Achemo_Brad == "Yes"                                 ~ last(sample_id)
#   # ), seqsample_id_Achemo = case_when(
#   #   seqsample_id_Achemo == sample_id                                ~ seqsample_id_Achemo,
#   #   TRUE                                                            ~ NA_character_
#   # )) %>% 
#   # ungroup() %>% 
#   # mutate(seqsample_date_Achemo = case_when(
#   #   !is.na(seqsample_id_Achemo)                                     ~ specimen_collection_date
#   # )) %>% 
#   # mutate(had_good_seqsample_Achemo = case_when(
#   #   seq_sample_Achemo_Brad == "Yes"                                 ~ "Yes"
#   # )) %>% 
#   # group_by(deidentified_patient_id) %>%
#   # mutate(after_chemo_seqsample_date = case_when(
#   #   had_good_seqsample_Achemo == "Yes"                              ~ (specimen_collection_date)
#   # )) %>% 
#   # # Look for seq sample after CR
#   # group_by(deidentified_patient_id) %>%
#   # fill(had_good_seqsample_Achemo, after_chemo_seqsample_date, .direction = "updown") %>%
#   # ungroup() %>%
#   # mutate(seq_sample_Achemo_Arad = case_when(
#   #   presample_chemo_rad_within_limits == "Yes" &
#   #     blood_bf_chemo_rad == "No" &
#   #     had_good_seqsample_Achemo == "Yes" &
#   #     specimen_collection_date > (radiation_start_date_1 + days(100)) &
#   #     specimen_collection_date <= (radiation_start_date_1 + days(365))  ~ "Yes",
#   #   TRUE                                                            ~ "No"
#   # )) %>%
#   # 
#   # group_by(mrn, seq_sample_Achemo_Arad) %>% 
#   # mutate(is_first_seq_sample = case_when(
#   #   seq_sample_Achemo_Arad == "Yes" &
#   #   specimen_collection_date == first(specimen_collection_date)     ~ "Yes",
#   #   TRUE                                                            ~ "No"
#   # )) %>% 
#   # ungroup() %>% 
#   # 
#   # mutate(seqsample_id_Achemorad = case_when(
#   #   is_first_seq_sample == "Yes"                                    ~ sample_id
#   # )) %>% 
#   # 
#   # mutate(seqsample_date_Achemorad = case_when(
#   #   !is.na(seqsample_id_Achemorad)                                  ~ specimen_collection_date
#   # )) %>% 
#   # mutate(had_good_seqsample_Achemorad = case_when(
#   #   seq_sample_Achemo_Arad == "Yes"                                 ~ "Yes"
#   # )) %>% 
#   # group_by(deidentified_patient_id) %>%
#   # fill(had_good_seqsample_Achemorad, .direction = "updown") %>%
#   # ungroup() %>%
#   
#   # mutate(time_chemo_1st_seqsample = case_when(
#   #   seq_sample_Achemo_Brad == "Yes"        ~ (
#   #     interval(start = chemotherapy_start_date_1, end = seqsample_date_Achemo) / 
#   #       duration(n= 1, units = "days")
#   # ))) %>% 
#   # mutate(time_1st_seqsample_rad = case_when(
#   #   seq_sample_Achemo_Brad == "Yes"        ~ (
#   #     interval(start = seqsample_date_Achemo, end = radiation_start_date_1) / 
#   #       duration(n= 1, units = "days")
#   # ))) %>% 
#   # mutate(time_rad_2nd_seqsample = case_when(
#   #   seq_sample_Achemo_Arad == "Yes"        ~ (
#   #     interval(start = radiation_start_date_1, end = seqsample_date_Achemorad) / 
#   #       duration(n= 1, units = "days")
#   # ))) %>% 
#   # mutate(interval_prepost_sample_chemo_rad = case_when(
#   #   !is.na(time_rad_2nd_seqsample)                                  ~ abs(
#   #     interval(start = after_chemo_seqsample_date, end = specimen_collection_date) /
#   #       duration(n = 1, units = "days")
#   # ))) %>% 
# blood_patients2 <- blood_patients2 %>%
#   mutate(across(where(is.numeric), ~ round(., 2)))

blood_patients2 <- read_rds(paste0(here::here(), "/blood_patients2.rds"))
```

# III. Patients with Blood before and after radiotharapy
## 1. Patients table 
```{r}
# tbl1_1 <- blood_patients2 %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(Chemotherapy = presample_chemo_within_limits) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
#   modify_header(update = list(
#   stat_0 ~ '**Patients with Blood before Treatment**'
# ))
# tbl2_1 <- blood_patients2 %>%
#     distinct(deidentified_patient_id, .keep_all = TRUE) %>%
#   select(Hormonetherapy = presample_hormone_within_limits) %>%
#   tbl_summary()
# tbl3 <- blood_patients2 %>% #filter(had_immuno == "Yes") %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(Immnunotherapy = blood_bf_immuno) %>% 
#   tbl_summary()
tbl4_1 <- blood_patients2 %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(Radiotherapy = presample_rad_within_limits) %>% 
  tbl_summary(
    statistic = list(all_categorical() ~ "{n}"),
    missing_text = "No"
  ) %>% 
  modify_header(update = list(
  stat_0 ~ '**Patients with Blood before Radiotherapy**'
))
# tbl5_1 <- blood_patients2 %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`Chemotherapy then Radiation` = presample_chemo_rad_within_limits) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl6_1 <- blood_patients2 %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`C+R or R+C` = presample_chemorad_within_limits) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl6 <- blood_patients2 %>% #filter(had_treatment == "Yes") %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`All Treatment` = blood_bf_treatment) %>% 
#   tbl_summary()

# tbl_s1 <- tbl_stack(list(tbl1, tbl4, tbl5))

# tbl1_2 <- blood_patients2 %>% 
#   filter(is.na(had_good_seqsample_Achemorad) & !is.na(seqsample_id_chemo)) %>%
#   
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(Chemotherapy = seq_sample_chemo) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
#   modify_header(update = list(
#   stat_0 ~ '**Patients with Sequential Blood after Treatment(s)**'
# ))
# tbl2_2 <- blood_patients2 %>% 
#   filter(!is.na(seqsample_id_hormone)) %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>%
#   select(Hormonetherapy = seq_sample_hormone) %>%
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl3 <- blood_patients2 %>% filter(had_immuno == "Yes") %>% 
#   filter(had_good_sample_chemo == "Yes") %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(Immnunotherapy = seq_sample_immuno) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
tbl4_2 <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_rad)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(Radiotherapy = seq_sample_rad) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
  modify_header(update = list(
  stat_0 ~ '**Patients with Sequential Blood after Radiotherapy**'
))
# tbl5_2 <- blood_patients2 %>% 
#   filter(!is.na(seqsample_id_Achemorad)) %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`Chemotherapy then Radiation` = had_good_seqsample_Achemorad) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl6_2 <- blood_patients2 %>%
#   filter(seq_sample_chemorad == "Yes") %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`C+R or R+C` = seq_sample_chemorad) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl6 <- blood_patients2 %>% #filter(had_treatment == "Yes") %>% 
#   filter(seq_sample_treatment == "Yes") %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`All Treatment` = seq_sample_treatment) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))

# tbl_s2 <- tbl_stack(list(tbl1, tbl4, tbl5#, tbl6
#                          ))

# tbl5_3 <- blood_patients2 %>% 
#   filter(!is.na(seqsample_id_Achemo)) %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`Chemotherapy then Radiation` = had_good_seqsample_Achemo) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
#   modify_header(update = list(
#   stat_0 ~ '**Patients with Sequential Blood after Chemo and before Radiation**'
# ))

# tbl_s1 <- tbl_merge(list(tbl1_1, tbl1_2))
# tbl_s2 <- tbl_merge(list(tbl2_1, tbl2_2))
tbl_s4 <- tbl_merge(list(tbl4_1, tbl4_2))
# tbl_s5 <- tbl_merge(list(tbl5_1, tbl5_2, tbl5_3))
# tbl_s6 <- tbl_merge(list(tbl6_1, tbl6_2))

tbl_s4 %>%
   modify_header(list(label ~ "**Treatment Type**")) %>% 
  modify_spanning_header(everything() ~ NA_character_)

# tbl_merge(list(tbl_s1, tbl_s2)) %>%
#   modify_spanning_header(everything() ~ NA_character_) %>% as_gt() %>%  
#   gt::tab_source_note(gt::md("**THe numbers reported show patients who have at least 1 sample before the corresponding treatment (`Yes') and patients from this selection with at least 1 sequential sample after treatment**"))

```

```{r}
rad_patients <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_rad))

rad_patients <- blood_patients2 %>%
  filter(
    str_detect(
      deidentified_patient_id, 
      paste0(rad_patients$deidentified_patient_id, collapse = "|")))
```

## 2. Time of samples before Radiotherapy
```{r}
sample_before_rad <- rad_patients %>% 
  filter(!is.na(interval_presample_rad)) %>%
  arrange(deidentified_patient_id, interval_presample_rad) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE)

sample_before_rad %>%
  ggplot(aes(x =interval_presample_rad
             ))+
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "purple") +
  theme_minimal(base_size = 14) +
  labs(x="Time from Blood Collection T1 to Radiation (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who had a sequential sample after Radiation")
```

## 3. Time of samples after Radiotherapy
```{r}
sample_after_rad <- rad_patients %>% 
  filter(!is.na(time_rad_seqsample)) %>%
  arrange(deidentified_patient_id, time_rad_seqsample) %>%
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>%
  group_by(deidentified_patient_id) %>%
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>%
  ungroup()

sample_after_rad %>%
  ggplot(aes(x =time_rad_seqsample, fill = sequential_sample_count))+
  geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Radiation start to \nT2 Sequential Blood Collection (in days)",
       y="Number of Patient",
       caption = "Each bar represents 100 days",
       title = "Patients who had a sequential sample after Radiation")

rad_patients %>%
  filter(!is.na(seqsample_date_rad)) %>%
  mutate(rad_start_T2 = interval(start = radiation_start_date_1, end = seqsample_date_rad)/
             duration(n=1, units = "days")) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select("Time from radiation start to T2" = rad_start_T2) %>% 
  tbl_summary() %>% as_gt() %>%
  gt::tab_source_note(gt::md("**Only for patients with stop date available**"))
```

## 4. Time between T1 and T2
```{r}
rad_patients %>% 
  filter(!is.na(time_rad_seqsample)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x =interval_prepost_sample_rad
             ))+
  geom_histogram(binwidth = 30, fill= "darkblue") +
  xlim(0, 400)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Pre Sample to \nPost Sample / Radiation (in days)", 
       y="Number of Patient",
       caption = "Each bar represents 30 days")
```

## 5. Did these patients receive drugs AFTER their first radiation? What drugs?
```{r}
rad_patients %>% 
  filter(!is.na(seqsample_date_rad)) %>%
  select(had_chemo, chemotherapy_drug_1) %>% 
  tbl_summary(missing_text = "Not administered")
```

## 6. Detailled table

```{r}
rad_patients <- rad_patients %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  mutate(ID = dense_rank(deidentified_patient_id))

df <- rad_patients %>%
  select(deidentified_patient_id, mrn, ID,
         "number post treatment sample"  = sequential_sample_count, 
         "days from pre sample to radiation" = interval_presample_rad,
         "days from radiation start to sequential sample" =  time_rad_seqsample)

color.me1 <- which((df$ID %% 2)  == 1)
# color.me <- which(df$`number post treatment sample` == 2)
# color.me.bold <- which((df$ID %% 2)  == 1)
color.me.pre <- which(df$`number post treatment sample` == 1)
color.me.seq1 <- which(df$`number post treatment sample` == 2)


df %>% select(-ID) %>% 
  kable(booktabs = T, align = "ccl") %>%
  # kable_styling(latex_options = "striped", stripe_color = "red") #%>%
  # row_spec(color.me, background = "#eee") %>%
  row_spec(color.me1, background = "#eee") %>%
  row_spec(color.me.pre, color = "blue") %>%
  row_spec(color.me.seq1, color = "red")

df1 <- df %>% filter(!is.na(`days from pre sample to radiation`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from pre sample to radiation`),2),
            median = round(median(`days from pre sample to radiation`),2))

df <- df %>% filter(!is.na(`days from radiation start to sequential sample`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from radiation start to sequential sample`),2),
            median = round(median(`days from radiation start to sequential sample`),2))
```
`days from pre sample to radiation` has a mean equal to `r df1$mean` and a median equal to `r df1$median`.  
`days from radiation start to sequential sample` has a mean equal to `r df$mean` and a median equal to `r df$median`.  
We have a list of `r length(unique(rad_patients$deidentified_patient_id))` unique patients.

<br>

## 7. Patients ids and age at diagnosis
```{r}
rad_patients %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(deidentified_patient_id, mrn, age_at_diagnosis)
```

## 8. Samples ids
```{r}
rad_patients %>%
  filter(!is.na(presample_id_rad) | !is.na(seqsample_id_rad)) %>% 
  select(deidentified_patient_id, mrn,
         presample_id_rad,
         presample_date_rad,
         seqsample_id_rad,
         seqsample_date_rad,
         time_rad_seqsample,
         radiation_start_date_1) %>%
  group_by(deidentified_patient_id) %>%
  fill(everything(), .direction = "updown") %>%
  distinct() %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%")
```

```{r}
rad <- rad_patients %>%
  filter(!is.na(presample_id_rad) | !is.na(seqsample_id_rad)) %>%
  select(mrn, deidentified_patient_id,
         presample_id_rad,
         presample_date_rad,
         interval_presample_rad,
         seqsample_id_rad,
         seqsample_date_rad,
         time_rad_seqsample,
         radiation_start_date_1) %>%
  group_by(deidentified_patient_id) %>%
  fill(everything(), .direction = "updown") %>%
  distinct()

write_csv(rad, "Samples list radiotherapy group in breast patients 01312023.csv")
```
<br>
