---
title: "Breast cancer and CH in Avatar data"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

<!-- table { -->
<!--   <!-- margin: auto; --> -->
<!--   border-top: 1px solid #666; -->
<!--   border-bottom: 1px solid #666; -->
<!-- } -->
<!-- table thead th { border-bottom: 1px solid #ddd; } -->
th, td { padding: 5px; }

<!-- tr:nth-of-type(3n+0) { -->
<!--   background: red; -->
<!-- } -->
<!-- tr:5:nth-child(0n+1) { -->
<!--   background: red; -->
<!-- } -->

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
library(ggforce)
library(kableExtra)
```

```{r}
# breast_dna <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/breast_dna.rds")
blood_patients <- 
  read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/blood_patients.rds")
# Treatment <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/Treatment.rds")
```

# Overall data selection
## Sample types selection
- collection_site_tissue_type == blood
- sample_type == Buffy Coat, Genomic DNA, MNC, MNC less CD138+, Unprocessed Liquid Tissue

## Patient characteristics
- primary_site_group == "Breast" (keep only breast diagnosis date)

## Treatment cleaning
- keep treatment for these selected patients (who has blood samples and breast cancer)
- remove drugs not used for breast cancer treatment
- removed drugs which were refused
- removed drugs that happened before the first date of diagnosis for breast cancer
- combined AC followed by P **when less than *not chosen* days between AC stop date and P start date**

# Repartition of Samples Before and After Treatments within Time Limits

We focus on the chemotherapy and radiation treatment so the samples are not chosen depending of immunotherapy or hormonetherapy. Also samples are chosen disregarding targeted therapy treatments.  

Summary criteria for sample selection:  
|•	Chemotherapy only:  
|-	Pre-sample must be closest to drug start from -365 days to + 5days   
|   also needs to be before rad or no rad was given  
|-	Sequential sample must be within a year from start date but after chemo stop date  
|   also needs to be before rad or no rad was given  

|•	Radiation only:  
|-	Pre-sample must be closest to drug start from -365 days to radiation date 
|   also needs to be before chemo or no chemo was given  
|-	Sequential sample must be within a year but 100 days after radiation  
|   also needs to be before chemo or no chemo was given  

•	Chemotherapy then Radiation:  
-	Pre-sample must be closest to chemo start from -365 days to + 5days   
-	Sequential sample T2  -> take all  
-	choose T2 closest to radiation start (if multiple samples)  
-	Sequential sample T3 must be within a year but 100 days after radiation  

|•	Hormonetherapy:  
|-	Pre-sample must be closest to drug start from -365 days to + 5days   
|   also needs to be before chemo/rad or no chemo/rad was given
|-	Sequential sample T2  -> take all  
|   also needs to be before chemo/rad or no chemo/rad was given

```{r}
blood_patients <- blood_patients %>% 
# blood_patients <- breast_dna %>% left_join(., Treatment, by = "mrn") %>% 
  mutate(blood_bf_chemo = case_when(
    specimen_collection_date <= (chemotherapy_start_date_1 + days(5)) &
      # specimen_collection_date <= specimen_collection_date - 365
      (specimen_collection_date <= radiation_start_date_1 |
      is.na(radiation_start_date_1))                                     ~ "Yes",
    specimen_collection_date > chemotherapy_start_date_1                 ~ "No",
    is.na(chemotherapy_start_date_1)                                     ~ "not administred",
    TRUE                                                                 ~ "No"
  )) %>% 
  mutate(blood_bf_hormone = case_when(
    specimen_collection_date <= (hormone_therapy_start_date_1 + days(5)) &
    (specimen_collection_date <= chemotherapy_start_date_1 |
      is.na(chemotherapy_start_date_1)) &
      (specimen_collection_date <= radiation_start_date_1 |
      is.na(radiation_start_date_1))                                        ~ "Yes",
    specimen_collection_date > hormone_therapy_start_date_1                 ~ "No",
    is.na(hormone_therapy_start_date_1)                                     ~ "not administred",
    TRUE                                                                    ~ "No"
  )) %>%
  # mutate(blood_bf_immuno = case_when(
  #   specimen_collection_date <= immunotherapy_start_date_1                ~ "Yes",
  #   specimen_collection_date > immunotherapy_start_date_1                 ~ "No",
  #   is.na(immunotherapy_start_date_1)                                     ~ "not administred",
  #   TRUE                                                                  ~ NA_character_
  # )) %>% 
  mutate(blood_bf_rad = case_when(
    specimen_collection_date <= radiation_start_date_1 &
      (specimen_collection_date <= chemotherapy_start_date_1 |
      is.na(chemotherapy_start_date_1))                                 ~ "Yes",
    specimen_collection_date > radiation_start_date_1                   ~ "No",
    is.na(radiation_start_date_1)                                       ~ "not administred",
    TRUE                                                                ~ "No"
  )) %>% 
  mutate(blood_bf_chemo_rad = case_when(
    specimen_collection_date <= (chemotherapy_start_date_1 + days(5)) &
      specimen_collection_date <= (radiation_start_date_1 + days(5))                ~ "Yes",
    specimen_collection_date > (chemotherapy_start_date_1 + days(5)) |
      specimen_collection_date > (radiation_start_date_1 + days(5))                 ~ "No",
    is.na(chemotherapy_start_date_1) |
      is.na(radiation_start_date_1)                                     ~ "both not administred",
    TRUE                                                                ~ "No"
  )) %>% 
  mutate(first_chemo_rad_date = case_when(
    chemotherapy_start_date_1 <= radiation_start_date_1                 ~ chemotherapy_start_date_1,
    chemotherapy_start_date_1 > radiation_start_date_1                  ~ radiation_start_date_1,
    TRUE                                                                ~ NA_Date_
  )) %>% 
  mutate(blood_bf_treatment = case_when(
    specimen_collection_date <= chemotherapy_start_date_1 &
      specimen_collection_date <= hormone_therapy_start_date_1 &
      specimen_collection_date <= immunotherapy_start_date_1 &
      specimen_collection_date <= radiation_start_date_1                ~ "Yes",
    if_any(contains("start_date_1"), ~ . < specimen_collection_date)    ~ "No",
    # if_all(contains("start_date_1"), ~ . < specimen_collection_date)    ~ "Nope",
    is.na(chemotherapy_start_date_1) |
      is.na(hormone_therapy_start_date_1) |
      is.na(immunotherapy_start_date_1) |
      is.na(radiation_start_date_1)                                     ~ "not administred",
    TRUE                                                                ~ "No"
  )) %>% 
  # mutate(blood_bf_30_days_chemo = case_when(
  #   specimen_collection_date >= (chemotherapy_start_date_1 - days(30)) &
  #     specimen_collection_date <= (chemotherapy_start_date_1 + days(30))              ~ "Yes",
  #   specimen_collection_date > (chemotherapy_start_date_1 + days(30))                 ~ "No",
  #   is.na(chemotherapy_start_date_1)                                     ~ "not administred",
  #   TRUE                                                            ~ NA_character_
  # )) %>%
  # mutate(blood_bf_30_days_hormone = case_when(
  #   specimen_collection_date <= (hormone_therapy_start_date_1 + days(30))                ~ "Yes",
  #   specimen_collection_date > (hormone_therapy_start_date_1 + days(30))                 ~ "No",
  #   is.na(hormone_therapy_start_date_1)                                     ~ "not administred",
  #   TRUE                                                            ~ NA_character_
  # )) %>%
  # mutate(blood_bf_30_days_immuno = case_when(
  #   specimen_collection_date <= (immunotherapy_start_date_1 + days(30))                ~ "Yes",
  #   specimen_collection_date > (immunotherapy_start_date_1 + days(30))                 ~ "No",
  #   is.na(immunotherapy_start_date_1)                                     ~ "not administred",
  #   TRUE                                                            ~ NA_character_
  # )) %>%
  # mutate(blood_bf_30_days_rad = case_when(
  #   specimen_collection_date <= (radiation_start_date_1 + days(30))                ~ "Yes",
  #   specimen_collection_date > (radiation_start_date_1 + days(30))                 ~ "No",
  #   is.na(radiation_start_date_1)                                     ~ "not administred",
  #   TRUE                                                            ~ NA_character_
  # )) %>%
  # mutate(blood_bf_30_days_chemo_rad = case_when(
  #   specimen_collection_date <= (chemotherapy_start_date_1 + days(30)) &
  #     specimen_collection_date <= (radiation_start_date_1 + days(30))                ~ "Yes",
  #   specimen_collection_date > (chemotherapy_start_date_1 + days(30)) &
  #     specimen_collection_date > (radiation_start_date_1 + days(30))                 ~ "No",
  #   is.na(chemotherapy_start_date_1) |
  #     is.na(radiation_start_date_1)                                 ~ "not administred",
  #   TRUE                                                            ~ NA_character_
  # )) %>%
  # mutate(blood_bf_30_days_treatment = case_when(
  #   specimen_collection_date <= (chemotherapy_start_date_1 + days(30)) &
  #     specimen_collection_date <= (hormone_therapy_start_date_1 + days(30)) &
  #     specimen_collection_date <= (immunotherapy_start_date_1 + days(30)) &
  #     specimen_collection_date <= (radiation_start_date_1 + days(30))                ~ "Yes",
  #   if_any(contains("start_date_1"), ~ (.  + days(30)) < specimen_collection_date)    ~ "No",
  #   # if_all(contains("start_date_1"), ~ . < specimen_collection_date)    ~ "Nope",
  #   is.na(chemotherapy_start_date_1) |
  #     is.na(hormone_therapy_start_date_1) |
  #     is.na(immunotherapy_start_date_1) |
  #     is.na(radiation_start_date_1)                                     ~ "not administred",
  #   TRUE                                                                ~ NA_character_
  # )) %>%
  mutate(across(contains("blood_bf_"), ~ factor(., levels = c("Yes", "No", "not administred")))) %>% 
  
  # ungroup() %>% 
  # mutate(treatment_bf_blood = case_when(
  #   specimen_collection_date <= treatment_start_date                ~ "Blood first",
  #   specimen_collection_date <= treatment_start_date                ~ treatment_line,
  #   TRUE                                                            ~ NA_character_
  # )) %>% 
  # mutate(treatment_bf_30days_blood = case_when(
  #   specimen_collection_date <= (treatment_start_date + days(30))   ~ "Blood first",
  #   specimen_collection_date <= (treatment_start_date + days(30))   ~ treatment_line,
  #   TRUE                                                            ~ NA_character_
  # )) %>% 
  # 
  # mutate(treatment_after_blood = case_when(
  #   specimen_collection_date > treatment_start_date                 ~ treatment_line,
  #   TRUE                                                            ~ NA_character_
  # )) %>% 
  # mutate(treatment_after_30days_blood = case_when(
  #   specimen_collection_date > (treatment_start_date + days(30))   ~ treatment_line,
  #   TRUE                                                            ~ NA_character_
  # )) %>% 
  # distinct(deidentified_patient_id, sample_family_id_sf, sample_id,
  #          specimen_collection_date, )
  arrange(mrn, deidentified_patient_id, specimen_collection_date, blood_bf_treatment) %>% 
  # group_by(mrn, deidentified_patient_id) %>% 
  # mutate(sample_lag = specimen_collection_date - lag(specimen_collection_date), 
  #        sample_lag = str_remove(sample_lag, " days")
  #        ) %>% 
  # mutate(has_a_good_sample = case_when(
  #   if_any(contains("blood_bf_"), ~ . == "Yes")    ~ "Yes"
  # )) %>% 
  # fill(has_a_good_sample, .direction = "down") %>% 
  # mutate(has_a_good_seq_sample = case_when(
  #   blood_bf_30_days_chemo_rad == "No" &
  #     has_a_good_sample == "Yes" &
  #     sample_lag > 30                             ~ "Yes"
  # )) %>% 
arrange(mrn, deidentified_patient_id) %>% 
  select(mrn, deidentified_patient_id, specimen_collection_date, #sample_lag, has_a_good_sample, has_a_good_seq_sample,
         "blood_bf_chemo", "blood_bf_rad",
         # "blood_bf_immuno", 
         "blood_bf_hormone", 
         "blood_bf_chemo_rad",
         "blood_bf_treatment", everything())
```


```{r code pre and sequential}
blood_patients2 <- blood_patients %>% 
  ###### Chemo samples ----
  mutate(had_good_sample_chemo = case_when(
    blood_bf_chemo == "Yes"                    ~ "Yes"
  )) %>%
  group_by(deidentified_patient_id, had_good_sample_chemo) %>%
  mutate(closest_chemo_presample_date = case_when(
    had_good_sample_chemo == "Yes"            ~ last(specimen_collection_date)
  )) %>%
  mutate(interval_presample_chemo = case_when(
    closest_chemo_presample_date == specimen_collection_date  ~ 
      (interval(start = closest_chemo_presample_date, end = chemotherapy_start_date_1) /
                 duration(n = 1, units = "days"))
  )) %>%
  mutate(presample_id_chemo = case_when(
    !is.na(interval_presample_chemo)       ~ sample_id
  )) %>% 
  mutate(presample_date_chemo = case_when(
    !is.na(presample_id_chemo)                                     ~ specimen_collection_date
  )) %>% 
  
  # Find sequential samples
  mutate(presample_chemo_within_limits = had_good_sample_chemo) %>% 
  group_by(deidentified_patient_id) %>%
  fill(presample_chemo_within_limits, #interval_presample_chemo,
       closest_chemo_presample_date, .direction = "updown") %>%
  mutate(seq_sample_chemo = case_when(
    presample_chemo_within_limits == "Yes" &
      blood_bf_chemo == "No" &
      specimen_collection_date > chemotherapy_end_date1_1 &
      specimen_collection_date <= (chemotherapy_start_date_1 + days(365)) &
      (specimen_collection_date <= radiation_start_date_1 |
      is.na(radiation_start_date_1))                                ~ "Yes",
    TRUE                                                            ~ "No"
  )) %>%
  mutate(seqsample_id_chemo = case_when(
    seq_sample_chemo == "Yes"                                       ~ sample_id
  )) %>% 
  mutate(seqsample_date_chemo = case_when(
    !is.na(seqsample_id_chemo)                                     ~ specimen_collection_date
  )) %>% 
  mutate(have_good_seqsample_chemo = case_when(
    seq_sample_chemo == "Yes"                                       ~ "Yes"
  )) %>% 
  fill(have_good_seqsample_chemo, .direction = "updown") %>% 
  mutate(time_chemo_seqsample = case_when(
    seq_sample_chemo == "Yes"        ~ (
      interval(start = chemotherapy_start_date_1, end = specimen_collection_date) /
        duration(n= 1, units = "days")
    ))) %>%
  mutate(interval_prepost_sample_chemo = case_when(
    !is.na(seqsample_id_chemo)                                     ~
      abs(interval(start = closest_chemo_presample_date, end = specimen_collection_date) /
                 duration(n = 1, units = "days"))
  )) %>%
  
  ###### Rad samples ----
  mutate(had_good_sample_rad = case_when(
    blood_bf_rad == "Yes"              ~ "Yes"
  )) %>%
  group_by(deidentified_patient_id, had_good_sample_rad) %>%
  mutate(closest_rad_presample_date = case_when(
    had_good_sample_rad == "Yes"       ~ last(specimen_collection_date)
  )) %>%
  mutate(interval_presample_rad = case_when(
    closest_rad_presample_date == specimen_collection_date  ~ 
      (interval(start = closest_rad_presample_date, end = radiation_start_date_1) /
                 duration(n = 1, units = "days"))
  )) %>%
  mutate(presample_id_rad = case_when(
    !is.na(interval_presample_rad)                           ~ sample_id
  )) %>% 
  mutate(presample_date_rad = case_when(
    !is.na(presample_id_rad)                                     ~ specimen_collection_date
  )) %>% 
  
  # Find sequential samples
  mutate(presample_rad_within_limits = had_good_sample_rad) %>%
  group_by(deidentified_patient_id) %>%
  fill(presample_rad_within_limits, #interval_presample_rad,
       closest_rad_presample_date, .direction = "updown") %>%
  mutate(seq_sample_rad = case_when(
    presample_rad_within_limits == "Yes" &
      blood_bf_rad == "No" &
      specimen_collection_date > (radiation_start_date_1 + days(100)) &
      specimen_collection_date <= (radiation_start_date_1 + days(365)) &
      (specimen_collection_date <= chemotherapy_start_date_1 |
      is.na(chemotherapy_start_date_1))                              ~ "Yes",
    TRUE ~ "No"
  )) %>%
  mutate(seqsample_id_rad = case_when(
    seq_sample_rad == "Yes"                                          ~ sample_id
  )) %>% 
  mutate(seqsample_date_rad = case_when(
    !is.na(seqsample_id_rad)                                     ~ specimen_collection_date
  )) %>% 
  mutate(have_good_seqsample_rad = case_when(
    seq_sample_rad == "Yes"                                       ~ "Yes"
  )) %>% 
  fill(have_good_seqsample_rad, .direction = "updown") %>% 
  mutate(time_rad_seqsample = case_when(
    seq_sample_rad == "Yes"        ~ (
      interval(start = radiation_start_date_1, end = specimen_collection_date) /
        duration(n= 1, units = "days")
  ))) %>%
  mutate(interval_prepost_sample_rad =
           abs(interval(start = closest_rad_presample_date, end = specimen_collection_date) /
                 duration(n = 1, units = "days"))) %>%

  ###### hormone samples ----
  mutate(had_good_sample_hormone = case_when(
    blood_bf_hormone == "Yes"              ~ "Yes"
  )) %>%
  group_by(deidentified_patient_id, had_good_sample_hormone) %>%
  mutate(closest_hormone_presample_date = case_when(
    had_good_sample_hormone == "Yes"       ~ last(specimen_collection_date)
  )) %>%
  mutate(interval_presample_hormone = case_when(
    closest_hormone_presample_date == specimen_collection_date  ~ 
      (interval(start = closest_hormone_presample_date, end = hormone_therapy_start_date_1) /
                 duration(n = 1, units = "days"))
  )) %>%
  mutate(presample_id_hormone = case_when(
    !is.na(interval_presample_hormone)                           ~ sample_id
  )) %>% 
  mutate(presample_date_hormone = case_when(
    !is.na(presample_id_hormone)                                     ~ specimen_collection_date
  )) %>% 
  
  # Find sequential samples
  mutate(presample_hormone_within_limits = had_good_sample_hormone) %>%
  group_by(deidentified_patient_id) %>%
  fill(presample_hormone_within_limits, #interval_presample_hormone,
       closest_hormone_presample_date, .direction = "updown") %>%
  mutate(seq_sample_hormone = case_when(
    presample_hormone_within_limits == "Yes" &
      blood_bf_hormone == "No" &
      specimen_collection_date > (hormone_therapy_start_date_1# + days(100)
                                  ) &
      # specimen_collection_date <= (hormone_therapy_start_date_1 + days(365)
      #                              ) &
      (specimen_collection_date <= chemotherapy_start_date_1 |
      is.na(chemotherapy_start_date_1)) &
      (specimen_collection_date <= radiation_start_date_1 |
      is.na(radiation_start_date_1))                                     ~ "Yes",
    TRUE ~ "No"
  )) %>%
  mutate(seqsample_id_hormone = case_when(
    seq_sample_hormone == "Yes"                                          ~ sample_id
  )) %>% 
  mutate(seqsample_date_hormone = case_when(
    !is.na(seqsample_id_hormone)                                     ~ specimen_collection_date
  )) %>% 
  mutate(have_good_seqsample_hormone = case_when(
    seq_sample_hormone == "Yes"                                       ~ "Yes"
  )) %>% 
  fill(have_good_seqsample_hormone, .direction = "updown") %>% 
  mutate(time_hormone_seqsample = case_when(
    seq_sample_hormone == "Yes"        ~ (
      interval(start = hormone_therapy_start_date_1, end = specimen_collection_date) /
        duration(n= 1, units = "days")
  ))) %>%
  mutate(interval_prepost_sample_hormone =
           abs(interval(start = closest_hormone_presample_date, end = specimen_collection_date) /
                 duration(n = 1, units = "days"))) %>%


  
  ###### C + Rad samples ---- 
  mutate(had_good_sample_chemrad = case_when(
    blood_bf_chemo_rad == "Yes"                              ~ "Yes"
  )) %>%
  group_by(deidentified_patient_id, had_good_sample_chemrad) %>%
  mutate(closest_chemorad_presample_date = case_when(
    had_good_sample_chemrad == "Yes"                         ~ last(specimen_collection_date)
  )) %>%
  mutate(interval_presample_chemrad = case_when(
    closest_chemorad_presample_date == specimen_collection_date  ~ 
      (interval(start = closest_chemorad_presample_date, end = first_chemo_rad_date) /
                 duration(n = 1, units = "days"))
  )) %>%
  mutate(presample_id_chemorad = case_when(
    !is.na(interval_presample_chemrad)                       ~ sample_id
  )) %>% 
  mutate(presample_date_chemorad = case_when(
    !is.na(presample_id_chemorad)                                     ~ specimen_collection_date
  )) %>% 
  
  # Find sequential samples
  mutate(presample_chemorad_within_limits = had_good_sample_chemrad) %>%
  group_by(deidentified_patient_id) %>%
  fill(presample_chemorad_within_limits, #interval_presample_chemrad,
       closest_chemorad_presample_date, .direction = "updown") %>%
  ungroup() %>% 
  # Intermediate var for seq sample
  mutate(interval_chemo_rad =
           abs(interval(start = chemotherapy_start_date_1, end = radiation_start_date_1) /
        duration(n= 1, units = "days"))
        ) %>%
  mutate(second_chemorad_date = first_chemo_rad_date + abs(interval_chemo_rad)) %>%
  mutate(seq_sample_chemorad = case_when(
    presample_chemorad_within_limits == "Yes" &
      blood_bf_chemo_rad == "No" &
      specimen_collection_date > (second_chemorad_date + days(100)) &
      specimen_collection_date <= (second_chemorad_date + days(365)) ~ "Yes",
    TRUE                                                     ~ "No"
  )) %>%
  mutate(seqsample_id_chemorad = case_when(
    seq_sample_chemorad == "Yes"                             ~ sample_id
  )) %>% 
  mutate(seqsample_date_chemorad = case_when(
    !is.na(seqsample_id_chemorad)                                     ~ specimen_collection_date
  )) %>% 
  mutate(have_good_seqsample_chemorad = case_when(
    seq_sample_chemorad == "Yes"                                       ~ "Yes"
  )) %>% 
  fill(have_good_seqsample_chemorad, .direction = "updown") %>% 
  # select(deidentified_patient_id, specimen_collection_date, had_good_sample_chemrad, closest_chemorad_presample_date, first_chemo_rad_date,second_chemrad_date,second_chemrad_date, seq_sample_chemorad, chemotherapy_start_date_1, radiation_start_date_1)
  mutate(time_chemorad_seqsample = case_when(
    seq_sample_chemorad == "Yes"                             ~ (
      interval(start = second_chemorad_date, end = specimen_collection_date) /
        duration(n= 1, units = "days")
  ))) %>%
  mutate(interval_prepost_sample_chemorad =
           abs(interval(start = closest_chemorad_presample_date, end = specimen_collection_date) /
                 duration(n = 1, units = "days"))) %>%

  ###### C then Rad sample ----
  mutate(had_good_sample_chemo_rad = case_when(
    blood_bf_chemo_rad == "Yes" &
      chemotherapy_start_date_1 < radiation_start_date_1     ~ "Yes"
  )) %>% 
  select(deidentified_patient_id, sample_id, specimen_collection_date, blood_bf_chemo_rad,
         chemotherapy_start_date_1, radiation_start_date_1, had_good_sample_chemo_rad, 
         first_chemo_rad_date, everything()) %>% 
  
  
  group_by(deidentified_patient_id, had_good_sample_chemo_rad) %>% 
  mutate(closest_chemo_rad_presample_date = case_when(
    had_good_sample_chemo_rad == "Yes"                       ~ last(specimen_collection_date)
  )) %>% 
  mutate(interval_presample_chemo_rad = case_when(
    closest_chemo_rad_presample_date == specimen_collection_date  ~ 
      (interval(start = closest_chemo_rad_presample_date, end = chemotherapy_start_date_1) /
                 duration(n = 1, units = "days"))
  )) %>%
  mutate(presample_id_chemo_rad = case_when(
    !is.na(interval_presample_chemo_rad)                     ~ sample_id
  )) %>% 
  mutate(presample_date_chemo_rad = case_when(
    !is.na(presample_id_chemo_rad)                                     ~ specimen_collection_date
  )) %>% 
  
  # Find seq sample after C and before R
  mutate(presample_chemo_rad_within_limits = had_good_sample_chemo_rad) %>%
  group_by(deidentified_patient_id) %>% 
  fill(presample_chemo_rad_within_limits, #closest_chemo_rad_presample_date, 
       .direction = "updown") %>% 
  ungroup() %>% 
  mutate(seq_sample_Achemo_Brad = case_when(
    presample_chemo_rad_within_limits == "Yes" &
      blood_bf_chemo_rad == "No" &
      # specimen_collection_date > (chemotherapy_start_date_1 + days(100)) &
      # specimen_collection_date > chemotherapy_end_date1_1 &
      specimen_collection_date > chemotherapy_start_date_1 &
      specimen_collection_date <= radiation_start_date_1            ~ "Yes",
    TRUE                                                            ~ "No"
  )) %>%
  group_by(deidentified_patient_id, seq_sample_Achemo_Brad) %>% 
  mutate(seqsample_id_Achemo = case_when(
    seq_sample_Achemo_Brad == "Yes"                                 ~ last(sample_id)
  ), seqsample_id_Achemo = case_when(
    seqsample_id_Achemo == sample_id                                ~ seqsample_id_Achemo,
    TRUE                                                            ~ NA_character_
  )) %>% 
  ungroup() %>% 
  mutate(seqsample_date_Achemo = case_when(
    !is.na(seqsample_id_Achemo)                                     ~ specimen_collection_date
  )) %>% 
  mutate(had_good_seqsample_Achemo = case_when(
    seq_sample_Achemo_Brad == "Yes"                                 ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>%
  mutate(after_chemo_seqsample_date = case_when(
    had_good_seqsample_Achemo == "Yes"                              ~ (specimen_collection_date)
  )) %>% 
  # Look for seq sample after CR
  group_by(deidentified_patient_id) %>%
  fill(had_good_seqsample_Achemo, after_chemo_seqsample_date, .direction = "updown") %>%
  ungroup() %>%
  mutate(seq_sample_Achemo_Arad = case_when(
    presample_chemo_rad_within_limits == "Yes" &
      blood_bf_chemo_rad == "No" &
      had_good_seqsample_Achemo == "Yes" &
      specimen_collection_date > (radiation_start_date_1 + days(100)) &
      specimen_collection_date <= (radiation_start_date_1 + days(365))  ~ "Yes",
    TRUE                                                            ~ "No"
  )) %>%
  mutate(seqsample_id_Achemorad = case_when(
    seq_sample_Achemo_Arad == "Yes"                                 ~ sample_id
  )) %>% 
  mutate(seqsample_date_Achemorad = case_when(
    !is.na(seqsample_id_Achemorad)                                  ~ specimen_collection_date
  )) %>% 
  mutate(had_good_seqsample_Achemorad = case_when(
    seq_sample_Achemo_Arad == "Yes"                                 ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>%
  fill(had_good_seqsample_Achemorad, .direction = "updown") %>%
  ungroup() %>%
  
  mutate(time_chemo_1st_seqsample = case_when(
    seq_sample_Achemo_Brad == "Yes"        ~ (
      interval(start = chemotherapy_start_date_1, end = seqsample_date_Achemo) / 
        duration(n= 1, units = "days")
  ))) %>% 
  mutate(time_1st_seqsample_rad = case_when(
    seq_sample_Achemo_Brad == "Yes"        ~ (
      interval(start = seqsample_date_Achemo, end = radiation_start_date_1) / 
        duration(n= 1, units = "days")
  ))) %>% 
  mutate(time_rad_2nd_seqsample = case_when(
    seq_sample_Achemo_Arad == "Yes"        ~ (
      interval(start = radiation_start_date_1, end = seqsample_date_Achemorad) / 
        duration(n= 1, units = "days")
  ))) %>% 
  mutate(interval_prepost_sample_chemo_rad = case_when(
    !is.na(time_rad_2nd_seqsample)                                  ~ abs(
      interval(start = after_chemo_seqsample_date, end = specimen_collection_date) /
        duration(n = 1, units = "days")
  ))) %>% 
  
  mutate(across(where(is.numeric), ~ round(., 2)))
           

  ###################
  # mutate(had_good_sample_treatment = case_when(
  #   blood_bf_treatment == "Yes"              ~ "Yes"
  # )) %>% 
  # group_by(deidentified_patient_id) %>% 
  # fill(had_good_sample_treatment, .direction = "updown") %>% 
  # mutate(seq_sample_treatment = case_when(
  #   had_good_sample_treatment == "Yes" &
  #     blood_bf_treatment == "No" ~ "Yes",
  #   TRUE ~ "No"
  # )) %>% 
  # ungroup() %>% 
  # select(deidentified_patient_id, specimen_collection_date, 
  #        blood_bf_chemo, closest_chemo_presample_date,
  #        "had_good_sample_chemo", "seq_sample_chemo",
  #        time_chemo_seqsample,
  #        everything())
```

# I. Data Exploration
## Missing stop date in selected patients?

```{r}
blood_patients2 %>% 
  filter(!is.na(seqsample_id_chemo)) %>% 
  mutate(missing_stop_date = case_when(
    is.na(chemotherapy_end_date1_1)      ~ "missing",
    TRUE                                 ~ "has chemo date"
  )) %>% 
  ggplot(aes(x= missing_stop_date)) +
  geom_bar()+
  theme_minimal(base_size = 14) +
  labs(y="Number of Patient",
       title = "Patients who have a pre AND sequential sample \n/ Chemotherapy")

blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad)) %>% 
  mutate(missing_stop_date = case_when(
    is.na(chemotherapy_end_date1_1)      ~ "missing",
    TRUE                                 ~ "has chemo date"
  )) %>% 
  ggplot(aes(x= missing_stop_date)) +
  geom_bar()+
  theme_minimal(base_size = 14) +
  labs(y="Number of Patient",
       title = "Patients who have a pre AND sequential sample \n/ Chemo-Rad")

blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>% 
  mutate(missing_stop_date = case_when(
    is.na(hormone_therapy_end_date1_1)   ~ "missing",
    TRUE                                 ~ "has hormone date"
  )) %>% 
  ggplot(aes(x= missing_stop_date)) +
  geom_bar()+
  theme_minimal(base_size = 14) +
  labs(y="Number of Patient",
       title = "Patients who have a pre AND sequential sample \n/ Hormonetherapy")
```
We have `r blood_patients2 %>% filter(!is.na(seqsample_id_hormone)) %>% filter(is.na(hormone_therapy_end_date1_1)) %>% nrow()` missing hormone stop date.

## What is the duration of the first regimen from start date to last stop date
### Chemo
```{r}
blood_patients2 %>% 
  filter(!is.na(seqsample_id_chemo)) %>% 
  mutate(regimen_duration = interval(start = chemotherapy_start_date_1, end = chemotherapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  ggplot(aes(x= regimen_duration)) +
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "darkgreen") +
  theme_minimal(base_size = 14) +
  labs(x="First regimen duration (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who have a pre AND sequential sample \n/ Chemotherapy")

blood_patients2 %>% 
  filter(!is.na(seqsample_id_chemo)) %>% 
  mutate(regimen_duration = interval(start = chemotherapy_start_date_1, end = chemotherapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  select(regimen_duration) %>% 
  tbl_summary()
blood_patients2 %>% 
  filter(!is.na(seqsample_id_chemo)) %>% 
  mutate(regimen_duration = interval(start = chemotherapy_start_date_1, end = chemotherapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  select(regimen_duration) %>% 
  tbl_summary(statistic =list(all_continuous() ~ "{mean} ({min}, {max})"))
```

### Chemo then Rad
```{r}
blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad)) %>% 
  mutate(regimen_duration = interval(start = chemotherapy_start_date_1, end = chemotherapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  ggplot(aes(x= regimen_duration)) +
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "darkgreen") +
  theme_minimal(base_size = 14) +
  labs(x="First regimen duration (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who have a pre AND sequential sample \n/ Chemo-Rad")

blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad)) %>% 
  mutate(regimen_duration = interval(start = chemotherapy_start_date_1, end = chemotherapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  select(regimen_duration) %>% 
  tbl_summary()
blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad)) %>% 
  mutate(regimen_duration = interval(start = chemotherapy_start_date_1, end = chemotherapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  select(regimen_duration) %>% 
  tbl_summary(statistic =list(all_continuous() ~ "{mean} ({min}, {max})"))
```
### Hormone
```{r}
blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>% 
  mutate(regimen_duration = interval(start = hormone_therapy_start_date_1, end = hormone_therapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  ggplot(aes(x= regimen_duration)) +
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "darkgreen") +
  theme_minimal(base_size = 14) +
  labs(x="First regimen duration (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who have a pre AND sequential sample \n/ Hormone")

blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>% 
  mutate(regimen_duration = interval(start = hormone_therapy_start_date_1, end = hormone_therapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  select(regimen_duration) %>% 
  tbl_summary()
blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>% 
  mutate(regimen_duration = interval(start = hormone_therapy_start_date_1, end = hormone_therapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  select(regimen_duration) %>% 
  tbl_summary(statistic =list(all_continuous() ~ "{mean} ({min}, {max})"))
```

# II. Table patients with Blood before and after treatment
```{r}
tbl1_1 <- blood_patients2 %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(Chemotherapy = presample_chemo_within_limits) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
  modify_header(update = list(
  stat_0 ~ '**Patients with Blood before Treatment**'
))
tbl2_1 <- blood_patients2 %>%
    distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(Hormonetherapy = presample_hormone_within_limits) %>%
  tbl_summary()
# tbl3 <- blood_patients2 %>% #filter(had_immuno == "Yes") %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(Immnunotherapy = blood_bf_immuno) %>% 
#   tbl_summary()
tbl4_1 <- blood_patients2 %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(Radiotherapy = presample_rad_within_limits) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
tbl5_1 <- blood_patients2 %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(`Chemotherapy then Radiation` = presample_chemo_rad_within_limits) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
tbl6_1 <- blood_patients2 %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(`C+R or R+C` = presample_chemorad_within_limits) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl6 <- blood_patients2 %>% #filter(had_treatment == "Yes") %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`All Treatment` = blood_bf_treatment) %>% 
#   tbl_summary()

# tbl_s1 <- tbl_stack(list(tbl1, tbl4, tbl5))

tbl1_2 <- blood_patients2 %>% 
  filter(is.na(seqsample_id_Achemorad) & !is.na(seqsample_id_chemo)) %>%
  
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(Chemotherapy = seq_sample_chemo) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
  modify_header(update = list(
  stat_0 ~ '**Patients with Sequential Blood after Treatment(s)**'
))
tbl2_2 <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>%
  select(Hormonetherapy = seq_sample_hormone) %>%
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl3 <- blood_patients2 %>% filter(had_immuno == "Yes") %>% 
#   filter(had_good_sample_chemo == "Yes") %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(Immnunotherapy = seq_sample_immuno) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
tbl4_2 <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_rad)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(Radiotherapy = seq_sample_rad) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
tbl5_2 <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(`Chemotherapy then Radiation` = had_good_seqsample_Achemorad) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
tbl6_2 <- blood_patients2 %>%
  filter(seq_sample_chemorad == "Yes") %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(`C+R or R+C` = seq_sample_chemorad) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl6 <- blood_patients2 %>% #filter(had_treatment == "Yes") %>% 
#   filter(seq_sample_treatment == "Yes") %>%
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(`All Treatment` = seq_sample_treatment) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))

# tbl_s2 <- tbl_stack(list(tbl1, tbl4, tbl5#, tbl6
#                          ))

tbl5_3 <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemo)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(`Chemotherapy then Radiation` = had_good_seqsample_Achemo) %>% 
  tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>% 
  modify_header(update = list(
  stat_0 ~ '**Patients with Sequential Blood after Chemo and before Radiation**'
))

tbl_s1 <- tbl_merge(list(tbl1_1, tbl1_2))
tbl_s2 <- tbl_merge(list(tbl2_1, tbl2_2))
tbl_s4 <- tbl_merge(list(tbl4_1, tbl4_2))
tbl_s5 <- tbl_merge(list(tbl5_1, tbl5_2, tbl5_3))
tbl_s6 <- tbl_merge(list(tbl6_1, tbl6_2))

tbl_stack(list(tbl_s1, tbl_s4, tbl_s5, tbl_s6, tbl_s2)) %>%
   modify_header(list(label ~ "**Treatment Type**")) %>% 
  modify_spanning_header(everything() ~ NA_character_) %>% as_gt() %>%  
  gt::tab_source_note(gt::md("*The number of patients with a blood sample before each treatment type is located 
                             in the first column. <br> 
                             From this number, the second column represent the number of patients with a blood 
                             sample after each treatment type. <br> 
                             For the case of Chemotherapy then Radiation, the number of patients with a blood
                             sample after Chemo and before Radiation is located in the third column. <br>
                             I added the 'C+R or R+C' treatment category in case we want extra samples after
                             both treatment for comparison. It represent a chemo and radiation treatment 
                             regardless of what comes first. I didn't look at the samples in between the 
                             treatment as it is not in the plan.*"))

# tbl_merge(list(tbl_s1, tbl_s2)) %>%
#   modify_spanning_header(everything() ~ NA_character_) %>% as_gt() %>%  
#   gt::tab_source_note(gt::md("**THe numbers reported show patients who have at least 1 sample before the corresponding treatment (`Yes') and patients from this selection with at least 1 sequential sample after treatment**"))

```

<!-- # Table total samples count before and after treatment (counting if multiple samples after) -->
```{r}
# tbl1 <- blood_patients2 %>% #filter(had_chemo == "Yes") %>% 
#   select(Chemotherapy = blood_bf_chemo) %>% 
#   tbl_summary() %>% 
#   modify_header(update = list(
#   stat_0 ~ '**Blood Samples before treatment**'
# ))
# 
# tbl2 <- blood_patients2 %>% #filter(had_hormone == "Yes") %>% 
#   select(Hormonetherapy = blood_bf_hormone) %>% 
#   tbl_summary()
# tbl3 <- blood_patients2 %>% #filter(had_immuno == "Yes") %>% 
#   select(Immnunotherapy = blood_bf_immuno) %>% 
#   tbl_summary()
# tbl4 <- blood_patients2 %>% #filter(had_rad == "Yes") %>% 
#   select(Radiotherapy = blood_bf_rad) %>% 
#   tbl_summary()
# tbl5 <- blood_patients2 %>% #filter(had_chemo_rad == "Yes") %>% 
#   select(`Chemotherapy+Radiation` = blood_bf_chemo_rad) %>% 
#   tbl_summary()
# tbl6 <- blood_patients2 %>% #filter(had_treatment == "Yes") %>% 
#   select(`All Treatment` = blood_bf_treatment) %>% 
#   tbl_summary()
# 
# tbl_s1 <- tbl_stack(list(tbl1, tbl2, tbl3, tbl4, tbl5, tbl6))
# 
# tbl1 <- blood_patients2 %>% filter(had_chemo == "Yes") %>% 
#   filter(seq_sample_chemo == "Yes") %>%
#   select(Chemotherapy = seq_sample_chemo) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}")) %>%
#   modify_header(update = list(
#   stat_0 ~ '**Sequential Blood Samples after Treatment**'
# ))
# # tbl2 <- blood_patients2 %>% filter(had_hormone == "Yes") %>% 
# #   filter(seq_sample_hormone == "Yes") %>%
# #   select(Hormonetherapy = seq_sample_hormone) %>% 
# #   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl3 <- blood_patients2 %>% filter(had_immuno == "Yes") %>%
#   filter(seq_sample_immuno == "Yes") %>%
#   select(Immnunotherapy = seq_sample_immuno) %>%
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl4 <- blood_patients2 %>% filter(had_rad == "Yes") %>% 
#   filter(seq_sample_rad == "Yes") %>%
#   select(Radiotherapy = seq_sample_rad) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# tbl5 <- blood_patients2 %>% filter(had_chemo_rad == "Yes") %>% 
#   filter(seq_sample_chemorad == "Yes") %>%
#   select(`Chemotherapy+Radiation` = seq_sample_chemorad) %>% 
#   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# # tbl6 <- blood_patients2 %>% #filter(had_treatment == "Yes") %>% 
# #   filter(seq_sample_treatment == "Yes") %>%
# #   select(`All Treatment` = seq_sample_treatment) %>% 
# #   tbl_summary(statistic = list(all_categorical() ~ "{n}"))
# 
# tbl_s2 <- tbl_stack(list(tbl1, tbl2, tbl3, tbl4, tbl5#, tbl6
#                          ))
# 
# tbl_merge(list(tbl_s1, tbl_s2)) %>%
#   modify_spanning_header(everything() ~ NA_character_) %>% as_gt() %>%  
#   gt::tab_source_note(gt::md("**THe numbers reported show total samples before the corresponding treatment (`Yes') and total sequential samples after treatment**"))

```

# III. Chemo group
```{r}
chemo_patients <- blood_patients2 %>% 
  filter(is.na(seqsample_id_Achemorad) & !is.na(seqsample_id_chemo))

chemo_patients <- blood_patients2 %>%
  filter(
    str_detect(
      deidentified_patient_id, 
      paste0(chemo_patients$deidentified_patient_id, collapse = "|")))
```

## Samples before treatment
```{r}
sample_before_chemo <- chemo_patients %>% 
  filter(!is.na(interval_presample_chemo)) %>%
  arrange(deidentified_patient_id, interval_presample_chemo) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE)

sample_before_chemo %>%
  ggplot(aes(x =interval_presample_chemo
             ))+
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "purple") +
  theme_minimal(base_size = 14) +
  labs(x="Time from Blood Collection T1 to Chemotherapy (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who had a sequential sample after Chemotherapy") +
  facet_zoom(xlim = c(-10, 150), zoom.size = 1)
```

## Samples after treatments
```{r}
sample_after_chemo <- chemo_patients %>% 
  filter(!is.na(time_chemo_seqsample)) %>%
  arrange(deidentified_patient_id, time_chemo_seqsample) %>%
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>%
  group_by(deidentified_patient_id) %>%
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>%
  ungroup()

sample_after_chemo %>%
  ggplot(aes(x =time_chemo_seqsample, fill = sequential_sample_count))+
  geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Chemotherapy start to \nT2 Sequential Blood Collection (in days)",
       y="Number of Patient",
       caption = "Each bar represents 100 days",
       title = "Patients who had a sequential sample after Chemotherapy")
```

## Time between chemotherapy stop date and T2

```{r}
chemo_patients %>%
  filter(!is.na(seqsample_date_chemo)) %>%
  mutate(chemo_stop_T2 = interval(start = chemotherapy_end_date1_1, end = seqsample_date_chemo)/
             duration(n=1, units = "days")) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x= chemo_stop_T2))+
  geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
  # scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Chemotherapy stop date to T2 (in days)", 
       y="Number of Patient",
       title = "Patients with sequential samples after Chemotherapy",
       caption = "Each bar represents 100 days")

chemo_patients %>%
  filter(!is.na(seqsample_date_chemo)) %>%
  mutate(chemo_stop_T2 = interval(start = chemotherapy_end_date1_1, end = seqsample_date_chemo)/
             duration(n=1, units = "days")) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(chemo_stop_T2) %>% 
  tbl_summary()
```

## Time between T1 and T2
```{r, fig.width=12, fig.height=9}
chemo_patients %>% 
  filter(!is.na(time_chemo_seqsample)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x =interval_prepost_sample_chemo
             ))+
  geom_histogram(binwidth = 30, fill= "darkblue") +
  theme_minimal(base_size = 14) +
  labs(x="Time from Pre Sample Chemotherapy to \nPost Sample (in days)", 
       y="Number of Patient",
       caption = "Each bar represents 30 days")

chemo_patients <- chemo_patients %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  mutate(ID = dense_rank(deidentified_patient_id))

df <- chemo_patients %>%
  select(deidentified_patient_id, ID,
         "number post treatment sample"  = sequential_sample_count, 
         "days from pre sample to chemo start" = interval_presample_chemo,
         "days from chemo start to sequential sample" =  time_chemo_seqsample)

color.me1 <- which((df$ID %% 2)  == 1)
# color.me <- which(df$`number post treatment sample` == 2)
# color.me.bold <- which((df$ID %% 2)  == 1)
color.me.pre <- which(df$`number post treatment sample` == 1)
color.me.seq1 <- which(df$`number post treatment sample` == 2)
```

I coded a de-identified ID for each patient, I used strips to visualize them better.  
This table only shows the samples after treatment and highlighted in <span style="color:red">red</span> is the first sample after treatment for each patient.
```{r}
df %>% select(-ID) %>% 
  kable(booktabs = T, align = "ccl") %>%
  # kable_styling(latex_options = "striped", stripe_color = "red") #%>%
  # row_spec(color.me, background = "#eee") %>%
  row_spec(color.me1, background = "#eee") %>%
  row_spec(color.me.pre, color = "blue") %>%
  row_spec(color.me.seq1, color = "red")

df1 <- df %>% filter(!is.na(`days from pre sample to chemo start`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from pre sample to chemo start`),2),
            median = round(median(`days from pre sample to chemo start`),2))

df <- df %>% filter(!is.na(`days from chemo start to sequential sample`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from chemo start to sequential sample`),2),
            median = round(median(`days from chemo start to sequential sample`),2))
```
`days from pre sample to chemo start` has a mean equal to `r df1$mean` and a median equal to `r df1$median`.  
`days from chemo start to sequential sample` has a mean equal to `r df$mean` and a median equal to `r df$median`.  
We have a list of `r length(unique(chemo_patients$deidentified_patient_id))` unique patients.

<br>

# List chemotherapy drugs (first regimen) 

```{r}
# blood_patients2 %>% 
#   filter(!is.na(seqsample_id_Achemorad) |
#            !is.na(seqsample_id_chemo)) %>% 
#   distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
#   select(chemotherapy_drug_1) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency"))

chemo_patients %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(chemotherapy_drug_1) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
```

# Patients list
```{r}
chemo_patients %>% 
  select(deidentified_patient_id, 
         presample_id_chemo, presample_date_chemo,
         seqsample_id_chemo, seqsample_date_chemo,
         time_chemo_seqsample,
         chemotherapy_start_date_1, chemotherapy_end_date1_1, 
         chemotherapy_drug_1) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(everything(), .direction = "updown") %>% 
  distinct() %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%")
```














# II. Sample intervals (focusing on patients who has a pre AND post samples)

## Samples before treatment

```{r}


sample_before_rad <- blood_patients2 %>%
  filter(!is.na(have_good_seqsample_rad)) %>%
  arrange(deidentified_patient_id, interval_presample_rad) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE)

sample_before_rad %>%
  ggplot(aes(x =interval_presample_rad
             ))+
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "purple") +
  theme_minimal(base_size = 14) +
  labs(x="Time from Blood Collection to Radiation (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who had a sequential sample after Radiation") +
  facet_zoom(xlim = c(0, 365), zoom.size = 1)

sample_before_chemo_rad <- blood_patients2 %>%
  filter(!is.na(had_good_seqsample_Achemorad)) %>%
  arrange(deidentified_patient_id, interval_presample_chemo_rad) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE)

sample_before_chemo %>%
  ggplot(aes(x =interval_presample_chemo_rad
             ))+
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "purple") +
  theme_minimal(base_size = 14) +
  labs(x="Time from Blood Collection to Chemotherapy (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Patients who had a sequential sample after Chemotherapy then Radiation") +
  facet_zoom(xlim = c(0, 365), zoom.size = 1)
```

## Samples after treatments

```{r Samples after treatments}


sample_after_hormone <- blood_patients2 %>%
  filter(!is.na(time_hormone_seqsample)) %>%
  arrange(deidentified_patient_id, time_hormone_seqsample) %>%
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>%
  group_by(deidentified_patient_id) %>%
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>%
  ungroup() %>%
  select(deidentified_patient_id, sequential_sample_count, time_hormone_seqsample)

sample_after_hormone %>%
  ggplot(aes(x =time_hormone_seqsample, fill = sequential_sample_count))+
  geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Hormonetherapy start to \nSequential Blood Collection (in days)",
       y="Number of Patient",
       caption = "Each bar represents 100 days",
       title = "Patients who had a sequential sample after Hormonetherapy")

sample_after_hormone %>%
  select(time_hormone_seqsample) %>% 
  tbl_summary()
sample_after_hormone %>%
  select(time_hormone_seqsample) %>% 
  tbl_summary(statistic =list(all_continuous() ~ "{mean} ({min}, {max})"))

blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>% 
  mutate(regimen_duration = interval(start = hormone_therapy_start_date_1, end = hormone_therapy_end_date1_1)/
           duration(n=1, units = "days")) %>% 
  ggplot(aes(x= regimen_duration)) +
  geom_histogram(binwidth = 7, alpha = 0.9, position = "stack", fill= "darkgreen") +
  theme_minimal(base_size = 14) +
  labs(x="First regimen duration (in days)",
       y="Number of Patient",
       caption = "Each bar represents 7 days",
       title = "Hormonetherapy") +
  facet_zoom(xlim = c(0, 1000), zoom.size = 1)
# sample_after_immuno <- blood_patients2 %>%
#   filter(seq_sample_immuno == "Yes") %>%
#   # mutate(interval_presample_immuno =
#   #          abs(interval(start = specimen_collection_date, end = immunotherapy_start_date_1) /
#   #                duration(n = 1, units = "days"))) %>%
#   arrange(deidentified_patient_id, interval_presample_immuno) %>%
#   distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   select(deidentified_patient_id, sequential_sample_count, interval_presample_immuno, everything())
# 
# sample_after_immuno %>% 
#   ggplot(aes(x =interval_presample_immuno, fill = sequential_sample_count))+
#   geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
#   scale_fill_viridis(discrete=T)+
#   theme_minimal(base_size = 14) +
#   labs(x="Time from Blood Collection to Immunotherapy (in days)", 
#        y="Number of Patient",
#        caption = "Each bar represents 100 days")

sample_after_rad <- blood_patients2 %>%
  filter(!is.na(time_rad_seqsample)) %>%
  arrange(deidentified_patient_id, time_rad_seqsample) %>%
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>%
  group_by(deidentified_patient_id) %>%
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>%
  ungroup()

sample_after_rad %>%
  ggplot(aes(x =time_rad_seqsample, fill = sequential_sample_count))+
  geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Radiotherapy to \nBlood Collection (in days)",
       y="Number of Patient",
       caption = "Each bar represents 100 days",
       title = "Patients who had a sequential sample after Radiotherapy")

sample_Achemo_Brad <- blood_patients2 %>%
  filter(!is.na(time_chemo_1st_seqsample)) %>%
  arrange(deidentified_patient_id, time_chemo_1st_seqsample) %>%
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  select(deidentified_patient_id, sequential_sample_count, time_chemo_1st_seqsample)

sample_Achemo_Brad %>% 
  ggplot(aes(x =time_chemo_1st_seqsample, fill = sequential_sample_count))+
  geom_histogram(binwidth = 30, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Chemotherapy start to \n2nd Sequential Sample (in days)", 
       y="Number of Patient",
       title = "Sequential samples after \nChemotherapy and before Radiotherapy",
       caption = "Each bar on the top plot represents 30 days") +
  ggforce::facet_zoom(xlim = c(0, 250), zoom.size = 2)


sample_Achemo_Arad <- blood_patients2 %>%
  filter(!is.na(time_rad_2nd_seqsample)) %>%
  arrange(deidentified_patient_id, time_rad_2nd_seqsample) %>%
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  select(deidentified_patient_id, sequential_sample_count, time_rad_2nd_seqsample)

sample_Achemo_Arad %>% 
  ggplot(aes(x =time_rad_2nd_seqsample, fill = sequential_sample_count))+
  geom_histogram(binwidth = 30, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Radiotherapy to \n 3rd Sequential Sample (in days)", 
       y="Number of Patient",
       title = "Sequential samples after \nChemotherapy and Radiotherapy",
       caption = "Each bar represents 30 days") +
  ggforce::facet_zoom(xlim = c(0, 270), zoom.size = 2)
```

## Time between Chemo and radiation for patients who had both sequential samples
```{r}
blood_patients2 %>%
  filter(!is.na(seqsample_id_Achemorad)) %>%
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x =interval_chemo_rad))+
  geom_histogram(binwidth = 100, alpha = 0.9, position = "stack") +
  # scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Chemotherapy to Radiotherapy (in days)", 
       y="Number of Patient",
       title = "Patients with sequential samples after Chemotherapy and Radiotherapy",
       caption = "Each bar represents 100 days")
```





# Time between samples and sample lists
I calculated the time between the closest pre treatment sample from treatment date (would "reference" sample would be a good term?) and the post treatment samples (I call them "sequential" samples)  

## Chemotherapy


<!-- ## Hormonetherapy -->
```{r, fig.width=12, fig.height=9}
# blood_patients2 %>% 
#   filter(!is.na(time_hormone_seqsample)) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   ggplot(aes(x =time_hormone_seqsample, fill = sequential_sample_count
#              ))+
#   geom_histogram(binwidth = 30, alpha = 0.9, position = "stack") +
#   scale_fill_viridis(discrete=T)+
#   theme_minimal(base_size = 14) +
#   labs(x="Time from Hormonetherapy Pre Treatment Sample to \nPost Treatment Sample (in days)", 
#        y="Number of Patient",
#        caption = "Each bar represents 30 days")
# 
# # knitr::kable(blood_patients2 %>% 
# #   filter(!is.na(time_prehormone_seq_sample)) %>% 
# #   group_by(deidentified_patient_id) %>% 
# #   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
# #   ungroup() %>% 
# #   mutate(ID = dense_rank(deidentified_patient_id)) %>% 
# #   select(ID, "number post treatment sample"  = sequential_sample_count, "days from closest pre treatment sample to post treatment sample" =  time_prehormone_seq_sample), align = "ccl")# %>%
#   # kable_styling() %>%
#   # cell_spec(bold = T, background = ifelse(ID > 6, "red", "white"))
# 
# df <- blood_patients2 %>% 
#   filter(!is.na(time_hormone_seqsample)) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   # mutate(ID = dense_rank(deidentified_patient_id)) %>% 
#   select(deidentified_patient_id, 
#          "days from pre sample to hormone" = interval_presample_hormone,
#          "number post treatment sample"  = sequential_sample_count, 
#          "days from treatment to sequential sample" =  time_hormone_seqsample)
# 
# color.me1 <- which((df$ID %% 2)  == 1)
# # color.me <- which(df$`number post treatment sample` == 2)
# # color.me.bold <- which((df$ID %% 2)  == 1)
# color.me.bold <- which(df$`number post treatment sample` == 1)
```

<!-- I recoded a dummy ID for each patient, I used strips to visualize them better.   -->
<!-- This table only shows the samples after treatment and highlighted in <span style="color:red">red</span> is the first sample after treatment for each patient. -->
```{r}
# df %>% 
#   kable(booktabs = T, align = "ccl") %>%
#   # kable_styling(latex_options = "striped", stripe_color = "red") #%>%
#   # row_spec(color.me, background = "#eee") %>%
#   row_spec(color.me1, background = "#eee") %>%
#   row_spec(color.me.bold, color = "red")
```
<br>

## Radiotherapy
```{r, fig.width=12, fig.height=9}
blood_patients2 %>% 
  filter(!is.na(time_rad_seqsample)) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  ggplot(aes(x =interval_prepost_sample_rad, fill = sequential_sample_count
             ))+
  geom_histogram(binwidth = 30, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Pre Sample Radiotherapy to \nPost Sample (in days)", 
       y="Number of Patient",
       caption = "Each bar represents 30 days")

# knitr::kable(blood_patients2 %>% 
#   filter(!is.na(time_prerad_seq_sample)) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   mutate(ID = dense_rank(deidentified_patient_id)) %>% 
#   select(ID, "number post treatment sample"  = sequential_sample_count, "days from closest pre treatment sample to post treatment sample" =  time_prerad_seq_sample), align = "ccl")# %>%
  # kable_styling() %>%
  # cell_spec(bold = T, background = ifelse(ID > 6, "red", "white"))

rad_patients <- blood_patients2 %>% 
  filter(!is.na(have_good_seqsample_rad)) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  mutate(ID = dense_rank(deidentified_patient_id))
df <- rad_patients %>%
  select(deidentified_patient_id, ID,
         "number post treatment sample"  = sequential_sample_count, 
         "days from pre sample to radiation" = interval_presample_rad,
         "days from radiation to sequential sample" =  time_rad_seqsample)

color.me1 <- which((df$ID %% 2)  == 1)
# color.me <- which(df$`number post treatment sample` == 2)
# color.me.bold <- which((df$ID %% 2)  == 1)
color.me.pre <- which(df$`number post treatment sample` == 1)
color.me.seq1 <- which(df$`number post treatment sample` == 2)
```


```{r table samples}
df %>% select(-ID) %>% 
  kable(booktabs = T, align = "ccl") %>%
  # kable_styling(latex_options = "striped", stripe_color = "red") #%>%
  # row_spec(color.me, background = "#eee") %>%
  row_spec(color.me1, background = "#eee") %>%
  row_spec(color.me.pre, color = "blue") %>%
  row_spec(color.me.seq1, color = "red")

df1 <- df %>% filter(!is.na(`days from pre sample to radiation`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from pre sample to radiation`),2),
            median = round(median(`days from pre sample to radiation`),2))

df <- df %>% filter(!is.na(`days from radiation to sequential sample`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from radiation to sequential sample`),2),
            median = round(median(`days from radiation to sequential sample`),2))
```
`days from pre sample to radiation` has a mean equal to `r df1$mean` and a median equal to `r df1$median`.  
`days from radiation to sequential sample` has a mean equal to `r df$mean` and a median equal to `r df$median`.  
We have a list of `r length(unique(rad_patients$deidentified_patient_id))` unique patients.

<br>


## Chemotherapy then Radiotherapy
```{r, fig.width=12, fig.height=9}
# blood_patients2 %>% 
#   filter(!is.na(had_good_seqsample_Achemorad)) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   ggplot(aes(x =time_chemo_1st_seqsample, fill = sequential_sample_count
#              ))+
#   geom_histogram(binwidth = 30, alpha = 0.9, position = "stack") +
#   scale_fill_viridis(discrete=T)+
#   theme_minimal(base_size = 14) +
#   labs(x="Time from Radiotherapy Pre Treatment Sample to \nPost Treatment Sample (in days)", 
#        y="Number of Patient",
#        caption = "Each bar represents 30 days")

# knitr::kable(blood_patients2 %>% 
#   filter(!is.na(time_prerad_seq_sample)) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   mutate(ID = dense_rank(deidentified_patient_id)) %>% 
#   select(ID, "number post treatment sample"  = sequential_sample_count, "days from closest pre treatment sample to post treatment sample" =  time_prerad_seq_sample), align = "ccl")# %>%
  # kable_styling() %>%
  # cell_spec(bold = T, background = ifelse(ID > 6, "red", "white"))

chemorad_patients <- blood_patients2 %>% 
  filter(!is.na(had_good_seqsample_Achemorad)) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  mutate(ID = dense_rank(deidentified_patient_id)) 
df <- chemorad_patients %>%
  select(deidentified_patient_id, ID,
         "number post treatment sample"  = sequential_sample_count, 
         "days from pre sample to chemo start" = interval_presample_chemo_rad,
         "days from chemo start to sequential sample (T2)"  = time_chemo_1st_seqsample,
         "days from sequential sample (T2) to radiation" = time_1st_seqsample_rad,
         "days from radiation to sequential sample (T3)" =  time_rad_2nd_seqsample,
         "days between T2 and T3"  = interval_prepost_sample_chemo_rad)

color.me1 <- which((df$ID %% 2)  == 1)
# color.me <- which(df$`number post treatment sample` == 2)
# color.me.bold <- which((df$ID %% 2)  == 1)
color.me.pre <- which(df$`number post treatment sample` == 1)
color.me.seq1 <- which(df$`number post treatment sample` == 2)
color.me.seq2 <- which(!is.na(df$`days from radiation to sequential sample (T3)`))
```


```{r}
df %>% select(-ID) %>% 
  kable(booktabs = T, align = "ccl") %>%
  # kable_styling(latex_options = "striped", stripe_color = "red") #%>%
  # row_spec(color.me, background = "#eee") %>%
  row_spec(color.me1, background = "#eee") %>%
  row_spec(color.me.pre, color = "blue") %>%
  row_spec(color.me.seq1, color = "red") %>%
  row_spec(color.me.seq2, color = "green")

df1 <- df %>% filter(!is.na(`days from pre sample to chemo start`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from pre sample to chemo start`),2),
            median = round(median(`days from pre sample to chemo start`),2))

df2 <- df %>% filter(!is.na(`days from chemo start to sequential sample (T2)`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from chemo start to sequential sample (T2)`),2),
            median = round(median(`days from chemo start to sequential sample (T2)`),2))

df <- df %>% filter(!is.na(`days from radiation to sequential sample (T3)`)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  summarize(mean = round(mean(`days from radiation to sequential sample (T3)`),2),
            median = round(median(`days from radiation to sequential sample (T3)`),2))
```
`days from pre sample to chemo start` has a mean equal to `r df1$mean` and a median equal to `r df1$median`.  
`days from chemo start to sequential sample (T2)` has a mean equal to `r df2$mean` and a median equal to `r df2$median`.  
`days from radiation to sequential sample (T3)` has a mean equal to `r df$mean` and a median equal to `r df$median`.  
We have a list of `r length(unique(chemorad_patients$deidentified_patient_id))` unique patients.

<br>

## Chemotherapy + Radiotherapy
```{r, fig.width=12, fig.height=9}
blood_patients2 %>% 
  filter(!is.na(time_chemorad_seqsample)) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  ggplot(aes(x =interval_prepost_sample_chemorad, fill = sequential_sample_count
             ))+
  geom_histogram(binwidth = 30, alpha = 0.9, position = "stack") +
  scale_fill_viridis(discrete=T)+
  theme_minimal(base_size = 14) +
  labs(x="Time from Pre Sample Chemo+Rad to \nPost Sample (in days)", 
       y="Number of Patient",
       caption = "Each bar represents 30 days")
# knitr::kable(blood_patients2 %>% 
#   filter(!is.na(time_prerad_seq_sample)) %>% 
#   group_by(deidentified_patient_id) %>% 
#   mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
#   ungroup() %>% 
#   mutate(ID = dense_rank(deidentified_patient_id)) %>% 
#   select(ID, "number post treatment sample"  = sequential_sample_count, "days from closest pre treatment sample to post treatment sample" =  time_prerad_seq_sample), align = "ccl")# %>%
  # kable_styling() %>%
  # cell_spec(bold = T, background = ifelse(ID > 6, "red", "white"))
df <- blood_patients2 %>% 
  filter(!is.na(time_chemorad_seqsample)) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sequential_sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  mutate(ID = dense_rank(deidentified_patient_id)) %>%
  select(deidentified_patient_id, ID,
         "days from pre sample to chemo or rad start" = interval_presample_chemrad,
         "number post treatment sample"  = sequential_sample_count, 
         "days from second treatment to sequential sample" =  time_chemorad_seqsample)
color.me1 <- which((df$ID %% 2)  == 1)
# color.me <- which(df$`number post treatment sample` == 2)
# color.me.bold <- which((df$ID %% 2)  == 1)
color.me.bold <- which(df$`number post treatment sample` == 1)
```


```{r}
df %>% select(-ID) %>% 
  kable(booktabs = T, align = "ccl") %>%
  # kable_styling(latex_options = "striped", stripe_color = "red") #%>%
  # row_spec(color.me, background = "#eee") %>%
  row_spec(color.me1, background = "#eee") %>%
  row_spec(color.me.bold, color = "red")
```
<br>

<!-- How patients overlap? -->
```{r}
# overlap <- 
#   full_join(chemorad_patients %>% 
#               select(deidentified_patient_id, presample_id_chemo_rad,
#                      seqsample_id_Achemo, seqsample_id_Achemorad),
#             chemo_patients %>% 
#               select(deidentified_patient_id, presample_id_chemo, 
#                      presample_id_chemo_rad,
#                      seqsample_id_chemo, 
#                      seqsample_id_Achemo), 
#             by = c("deidentified_patient_id", "presample_id_chemo_rad", "seqsample_id_Achemo")) %>% 
#   full_join(.,
#             rad_patients %>% 
#               select(deidentified_patient_id, presample_id_rad,
#                      seqsample_id_rad),
#             by = "deidentified_patient_id") %>% 
#   arrange(presample_id_chemo_rad, presample_id_chemo) %>% 
#   mutate(common_presample = case_when(
#     presample_id_chemo == presample_id_chemo_rad             ~ "Yes",
#     presample_id_chemo != presample_id_chemo_rad             ~ "No"
#   )) %>% 
#   mutate(common_seqsample_chemo = case_when(
#     presample_id_chemo == presample_id_chemo_rad             ~ "Yes",
#     presample_id_chemo != presample_id_chemo_rad             ~ "No"
#   )) %>% 
#   arrange(deidentified_patient_id)

```


# List chemotherapy drugs (first regimen) 

```{r}
blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad) |
           !is.na(seqsample_id_chemo)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(chemotherapy_drug_1) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))

blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(hormone_therapy_drug_1) %>% 
  tbl_summary(sort = list(everything() ~ "frequency"))
```

# Patients list
## Chemo then Rad
```{r}
chemorad_patients <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_Achemorad)) 
blood_patients2 %>% 
  filter(str_detect(deidentified_patient_id,
                    paste0(chemorad_patients$deidentified_patient_id, 
                           collapse = "|"))) %>% 
  select(deidentified_patient_id, 
         presample_id_chemorad, presample_date_chemorad,
         seqsample_id_Achemo, seqsample_date_Achemo,
         seqsample_id_Achemorad, seqsample_date_Achemorad,
         "days from radiation to sequential sample (T3)" =  time_rad_2nd_seqsample,
         chemotherapy_start_date_1, chemotherapy_end_date1_1, 
         chemotherapy_drug_1, radiation_start_date_1) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(everything(), .direction = "updown") %>% 
  distinct() %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%")
```



## Rad
```{r}
rad_patients <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_rad)) 
blood_patients2 %>% 
  filter(str_detect(deidentified_patient_id,
                    paste0(rad_patients$deidentified_patient_id, 
                           collapse = "|"))) %>% 
  select(deidentified_patient_id, 
         presample_id_rad, presample_date_rad,
         seqsample_id_rad, seqsample_date_rad,
         time_rad_seqsample,
         radiation_start_date_1) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(everything(), .direction = "updown") %>% 
  distinct() %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%")
```

# Hormone
```{r}
hormone_patients <- blood_patients2 %>% 
  filter(!is.na(seqsample_id_hormone)) 
blood_patients2 %>% 
  filter(str_detect(deidentified_patient_id,
                    paste0(hormone_patients$deidentified_patient_id, 
                           collapse = "|"))) %>% 
  select(deidentified_patient_id, 
         presample_id_hormone, presample_date_hormone,
         seqsample_id_hormone, seqsample_date_hormone,
         time_hormone_seqsample,
         hormone_therapy_start_date_1, hormone_therapy_end_date1_1,
         hormone_therapy_drug_1) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(everything(), .direction = "updown") %>% 
  distinct() %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%")
```

# ER status

```{r}
# blood_patients2 %>% 
#   filter(str_detect(deidentified_patient_id,
#                     paste0(chemorad_patients$deidentified_patient_id, 
#                            collapse = "|"))) %>% 
#   select(ER)
```


## Let me know if you need anything else!

***











