---
title: "Secondary Biagnosis Breast cancer"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
```

```{r}
Second_dx <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/Second_dx.rds")
# breast_dna <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/breast_dna.rds")
# Treatment <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/Treatment.rds")
```

# How many patients in the breast cancer cohort that you have had a secondary diagnosis of a myeloid malignancy?
## General Cancer type

I selected all the cancer which contains the word "BREAST" in the cancer_site variable as breast cancer.  
Please could you let me know if I forgot others. All cancer sites present in the data are in the table below.  

```{r}
Second_dx %>% 
  select(cancer_site) %>% 
  tbl_summary()
```

Here is a summary of the raw histology variable if it can help

```{r}
Second_dx %>% 
  select(histology) %>% 
  tbl_summary()
```

If it is no more cancer_site to add as being breast cancer. I categorize the other type of cancer as "pre" and "post" breast cancer. I think you are only interested by the secondary cancer but I thought I will put it that out there.
I also had 2 patients who had a diagnose for another cancer on the same date. The last category is the relapse for breast cancer (this category will disappear when we focus on myeloid malignancy `r emo::ji("smiling_face_with_smiling_eyes")`).  

```{r}
Second_dx <- Second_dx %>% 
  # Create a variable for the first breast diagnose for each patient
  arrange(deidentified_patient_id, date_of_diagnosis) %>% 
  mutate(breast_cancer_dx = case_when(
    str_detect(cancer_site, "BREAST")                  ~ date_of_diagnosis
  )) %>% 
  select(deidentified_patient_id, breast_cancer_dx, everything()) %>% 
  arrange(deidentified_patient_id, breast_cancer_dx) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(breast_cancer_dx = first(breast_cancer_dx)) %>% 
  ungroup() %>% 
  # Remove patients who never had breast cancer
  filter(!is.na(breast_cancer_dx)) %>% 
  # was the primary or secondary cancer
  mutate(relative_cancer = case_when(
    str_detect(cancer_site, "BREAST") &
      breast_cancer_dx == date_of_diagnosis             ~ "breast",
    str_detect(cancer_site, "BREAST")                   ~ "relapse",
    breast_cancer_dx < date_of_diagnosis                ~ "post-breast cancer",
    breast_cancer_dx > date_of_diagnosis                ~ "pre-breast cancer",
    breast_cancer_dx == date_of_diagnosis               ~ "diagnosed at the same time"
  ))

Second_dx %>% 
  filter(relative_cancer != "breast") %>% 
  select(cancer_site, relative_cancer) %>% 
  tbl_summary(by = relative_cancer)

Second_dx %>% 
  filter(relative_cancer != "breast") %>% 
  select(histology, relative_cancer) %>% 
  tbl_summary(by = relative_cancer)
```

Please let me know which diagnoses I can include to look at the sample availability.  










