---
title: "Secondary Biagnosis Breast cancer"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(lubridate)
library(gtsummary)
library(viridis)
```

```{r}
Second_dx <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/Second_dx.rds")
breast_dna <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/breast_dna.rds")
Treatment <- read_rds("/Users/colinccm/Documents/GitHub/Gillis/breast_ch/Treatment.rds")
```

# How many patients in the breast cancer cohort that you have had a secondary diagnosis of a myeloid malignancy?
## General Cancer type

I selected all the cancer which contains the word "BREAST" in the cancer_site variable as breast cancer.  
Please could you let me know if I forgot others. All cancer sites present in the data are in the table below.  

```{r}
Second_dx %>% 
  select(cancer_site) %>% 
  tbl_summary()
```

If it is no more cancer_site to add as being breast cancer. I categorize the other type of cancer as "pre" and "post" breast cancer. I think you are only interested by the secondary cancer but I thought I will put it that out there.
I also had 2 patients who had a diagnose for another cancer on the same date. The last category is the relapse for breast cancer (this category will disappear when we focus on myeloid malignancy `r emo::ji("smiling_face_with_smiling_eyes")`).  

```{r}
Second_dx <- Second_dx %>% 
  # Create a variable for the first breast diagnose for each patient
  arrange(deidentified_patient_id, date_of_diagnosis) %>% 
  mutate(breast_cancer_dx = case_when(
    str_detect(cancer_site, "BREAST")                  ~ date_of_diagnosis
  )) %>% 
  select(deidentified_patient_id, breast_cancer_dx, everything()) %>% 
  arrange(deidentified_patient_id, breast_cancer_dx) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(breast_cancer_dx = first(breast_cancer_dx)) %>% 
  ungroup() %>% 
  # Remove patients who never had breast cancer
  filter(!is.na(breast_cancer_dx)) %>% 
  # was the primary or secondary cancer
  mutate(relative_cancer = case_when(
    str_detect(cancer_site, "BREAST") &
      breast_cancer_dx == date_of_diagnosis             ~ "breast",
    str_detect(cancer_site, "BREAST")                   ~ "relapse",
    breast_cancer_dx < date_of_diagnosis                ~ "post-breast cancer",
    breast_cancer_dx > date_of_diagnosis                ~ "pre-breast cancer",
    breast_cancer_dx == date_of_diagnosis               ~ "diagnosed at the same time"
  ))

Second_dx %>% 
  filter(relative_cancer != "breast") %>% 
  select(cancer_site, relative_cancer) %>% 
  tbl_summary(by = relative_cancer)
```

Please let me know which diagnoses I can include to look at the sample availability.  

# How many sarcoma patients have samples from more than one time/date?

```{r}
breast_dna1 <- breast_dna %>% left_join(., Treatment, by = "deidentified_patient_id") %>% 
  mutate(blood_bf_chemo = case_when(
    specimen_collection_date <= chemotherapy_start_date_1                ~ "Yes",
    specimen_collection_date > chemotherapy_start_date_1                 ~ "No",
    is.na(chemotherapy_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>% 
  mutate(blood_bf_hormone = case_when(
    specimen_collection_date <= hormone_therapy_start_date_1                ~ "Yes",
    specimen_collection_date > hormone_therapy_start_date_1                 ~ "No",
    is.na(hormone_therapy_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>% 
  mutate(blood_bf_immuno = case_when(
    specimen_collection_date <= immunotherapy_start_date_1                ~ "Yes",
    specimen_collection_date > immunotherapy_start_date_1                 ~ "No",
    is.na(immunotherapy_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>% 
  mutate(blood_bf_rad = case_when(
    specimen_collection_date <= radiation_start_date_1                ~ "Yes",
    specimen_collection_date > radiation_start_date_1                 ~ "No",
    is.na(radiation_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>% 
  mutate(blood_bf_chemo_rad = case_when(
    specimen_collection_date <= chemotherapy_start_date_1 &
      specimen_collection_date <= radiation_start_date_1                ~ "Yes",
    specimen_collection_date > chemotherapy_start_date_1 |
      specimen_collection_date > radiation_start_date_1                 ~ "No",
    is.na(chemotherapy_start_date_1) |
      is.na(radiation_start_date_1)                                 ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>% 
  mutate(blood_bf_treatment = case_when(
    specimen_collection_date <= chemotherapy_start_date_1 &
      specimen_collection_date <= hormone_therapy_start_date_1 &
      specimen_collection_date <= immunotherapy_start_date_1 &
      specimen_collection_date <= radiation_start_date_1                ~ "Yes",
    if_any(contains("start_date_1"), ~ . < specimen_collection_date)    ~ "No",
    # if_all(contains("start_date_1"), ~ . < specimen_collection_date)    ~ "Nope",
    is.na(chemotherapy_start_date_1) |
      is.na(hormone_therapy_start_date_1) |
      is.na(immunotherapy_start_date_1) |
      is.na(radiation_start_date_1)                                     ~ "not administred",
    TRUE                                                                ~ NA_character_
  )) %>% 
  mutate(blood_bf_30_days_chemo = case_when(
    specimen_collection_date >= (chemotherapy_start_date_1 - days(30)) &
      specimen_collection_date <= (chemotherapy_start_date_1 + days(30))              ~ "Yes",
    specimen_collection_date > (chemotherapy_start_date_1 + days(30))                 ~ "No",
    is.na(chemotherapy_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>%
  mutate(blood_bf_30_days_hormone = case_when(
    specimen_collection_date <= (hormone_therapy_start_date_1 + days(30))                ~ "Yes",
    specimen_collection_date > (hormone_therapy_start_date_1 + days(30))                 ~ "No",
    is.na(hormone_therapy_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>%
  mutate(blood_bf_30_days_immuno = case_when(
    specimen_collection_date <= (immunotherapy_start_date_1 + days(30))                ~ "Yes",
    specimen_collection_date > (immunotherapy_start_date_1 + days(30))                 ~ "No",
    is.na(immunotherapy_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>%
  mutate(blood_bf_30_days_rad = case_when(
    specimen_collection_date <= (radiation_start_date_1 + days(30))                ~ "Yes",
    specimen_collection_date > (radiation_start_date_1 + days(30))                 ~ "No",
    is.na(radiation_start_date_1)                                     ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>%
  mutate(blood_bf_30_days_chemo_rad = case_when(
    specimen_collection_date <= (chemotherapy_start_date_1 + days(30)) &
      specimen_collection_date <= (radiation_start_date_1 + days(30))                ~ "Yes",
    specimen_collection_date > (chemotherapy_start_date_1 + days(30)) &
      specimen_collection_date > (radiation_start_date_1 + days(30))                 ~ "No",
    is.na(chemotherapy_start_date_1) |
      is.na(radiation_start_date_1)                                 ~ "not administred",
    TRUE                                                            ~ NA_character_
  )) %>%
  mutate(blood_bf_30_days_treatment = case_when(
    specimen_collection_date <= (chemotherapy_start_date_1 + days(30)) &
      specimen_collection_date <= (hormone_therapy_start_date_1 + days(30)) &
      specimen_collection_date <= (immunotherapy_start_date_1 + days(30)) &
      specimen_collection_date <= (radiation_start_date_1 + days(30))                ~ "Yes",
    if_any(contains("start_date_1"), ~ (.  + days(30)) < specimen_collection_date)    ~ "No",
    # if_all(contains("start_date_1"), ~ . < specimen_collection_date)    ~ "Nope",
    is.na(chemotherapy_start_date_1) |
      is.na(hormone_therapy_start_date_1) |
      is.na(immunotherapy_start_date_1) |
      is.na(radiation_start_date_1)                                     ~ "not administred",
    TRUE                                                                ~ NA_character_
  )) %>%
  mutate(across(contains("blood_bf_"), ~ factor(., levels = c("Yes", "No", "not administred")))) %>% 
  
  arrange(deidentified_patient_id, specimen_collection_date, blood_bf_treatment) %>% 
  group_by(deidentified_patient_id) %>% 

  select(deidentified_patient_id, specimen_collection_date, #sample_lag, has_a_good_sample, has_a_good_seq_sample,
         "blood_bf_chemo", "blood_bf_hormone", 
         "blood_bf_immuno", "blood_bf_rad",
         "blood_bf_chemo_rad",
         "blood_bf_treatment", everything())

breast_dna2 <- breast_dna1 %>% 
  mutate(had_good_sample_chemo = case_when(
    blood_bf_chemo == "Yes"              ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(had_good_sample_chemo, .direction = "updown") %>% 
  mutate(seq_sample_chemo = case_when(
    had_good_sample_chemo == "Yes" &
      blood_bf_chemo == "No" ~ "Yes",
    TRUE ~ "No"
  )) %>% 
  mutate(had_good_sample_hormone = case_when(
    blood_bf_hormone == "Yes"              ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(had_good_sample_hormone, .direction = "updown") %>% 
  mutate(seq_sample_hormone = case_when(
    had_good_sample_hormone == "Yes" &
      blood_bf_hormone == "No" ~ "Yes",
    TRUE ~ "No"
  )) %>% 
  mutate(had_good_sample_immuno = case_when(
    blood_bf_immuno == "Yes"              ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(had_good_sample_immuno, .direction = "updown") %>% 
  mutate(seq_sample_immuno = case_when(
    had_good_sample_immuno == "Yes" &
      blood_bf_immuno == "No" ~ "Yes",
    TRUE ~ "No"
  )) %>% 
  mutate(had_good_sample_rad = case_when(
    blood_bf_rad == "Yes"              ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(had_good_sample_rad, .direction = "updown") %>% 
  mutate(seq_sample_rad = case_when(
    had_good_sample_rad == "Yes" &
      blood_bf_rad == "No" ~ "Yes",
    TRUE ~ "No"
  )) %>% 
  mutate(had_good_sample_chemo_rad = case_when(
    blood_bf_chemo_rad == "Yes"              ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(had_good_sample_chemo_rad, .direction = "updown") %>% 
  mutate(seq_sample_chemo_rad = case_when(
    had_good_sample_chemo_rad == "Yes" &
      blood_bf_chemo_rad == "No" ~ "Yes",
    TRUE ~ "No"
  )) %>% 
  mutate(had_good_sample_treatment = case_when(
    blood_bf_treatment == "Yes"              ~ "Yes"
  )) %>% 
  group_by(deidentified_patient_id) %>% 
  fill(had_good_sample_treatment, .direction = "updown") %>% 
  mutate(seq_sample_treatment = case_when(
    had_good_sample_treatment == "Yes" &
      blood_bf_treatment == "No" ~ "Yes",
    TRUE ~ "No"
  )) %>% 
  ungroup()
```

# Number of samples from more than one date
```{r}
breast_dna2 %>% 
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_count = factor(row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  arrange(desc(sample_count)) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  select(sample_count) %>% 
  tbl_summary()
```

I don't think you are interested about when the sample was collected (before or after chemotherapy) but let me know if you are.  

# Histology summary

```{r}
breast_dna2 %>% 
  distinct(deidentified_patient_id, specimen_collection_date, .keep_all = TRUE) %>% 
  group_by(deidentified_patient_id) %>% 
  mutate(sample_count = (row_number(deidentified_patient_id))) %>% 
  ungroup() %>% 
  filter(sample_count > 1) %>% 
  distinct(deidentified_patient_id, .keep_all = TRUE) %>% 
  left_join(., Second_dx %>% select(deidentified_patient_id, histology), 
            by = "deidentified_patient_id") %>% 
  select(deidentified_patient_id, histology) %>% 
    tbl_summary()









```








